% ================================================================
% ==== FILE: content/predictions/predictions_k0.tex
% ================================================================

% ==============================
%  Ontology of Continua — Core
%  Predictions for K0 — FULL VERSION
% ==============================

\subsubsection{Predictions for \texorpdfstring{$K_0$}{K_0}}
\label{sec:predictions-k0}

Predictions at level $K_0$ follow from the minimal structure  
\[
K_0 = (S, \Delta, \mathcal{C})
\]
equipped with the axioms of locality, regularity, admissible compositions  
and the minimal existence threshold $\Theta_0 = \varepsilon > 0$.

Because $K_0$ contains no axes, no potentials, no flows and no cycles,  
all predictions are structural and concern the necessary behaviour of  
\emph{any} proto-continuum that may give rise to higher levels $K_1, K_2,\dots$.

The predictions of $K_0$ fall into five groups.

% ---------------------------------------------------------------
\subsubsection{P1: Predictions Concerning Existence}

\paragraph{(P1.1) Minimal Extent Prediction.}
If a proto-continuum exists, its configuration set $\Delta$ must satisfy
\[
|\Delta| \ge 1,
\]
otherwise no continuum exists, even in degenerate form.

\paragraph{(P1.2) Non-zero Structural Threshold.}
The minimal threshold
\[
\Theta_0 = \varepsilon > 0
\]
predicts that no continuum can exist with exact zero structural slack.  
Any structure with $\Theta_0 = 0$ is forbidden:
\[
\Theta_0 = 0 \quad\Rightarrow\quad K_0 \text{ impossible}.
\]

\paragraph{(P1.3) No Perfect Determinacy.}
Because admissible configurations form $\Delta$ rather than a single element,
\[
|\Delta| = 1 \Rightarrow K_0 \text{ collapses},
\]
i.e. a proto-continuum must have at least some configurational latitude.

% ---------------------------------------------------------------
\subsubsection{P2: Predictions Concerning Structure and Composition}

\paragraph{(P2.1) Compositional Closure Prediction.}
From axioms of $\mathcal{C}$:
\[
\mathcal{C}(x,y) \text{ defined} \Rightarrow \mathcal{C}(y,x) \text{ not required}.
\]
Thus OC predicts that $K_0$ composition is not necessarily symmetric,  
supporting the later emergence of directional dynamics ($K_1$).

\paragraph{(P2.2) No Global Composition Prediction.}
$K_0$ predicts:
\[
\neg \exists\, \mathcal{C}_{\mathrm{global}} : S\times S\to S.
\]
Global closure is impossible; only partial compositions are admissible.

This matches the observed emergence of local-to-global structure only at $K_1$.

\paragraph{(P2.3) Partiality is Necessary, not Contingent.}
A full composition table is forbidden:
\[
\forall x,y\in S,\ \mathcal{C}(x,y)\ \text{defined}
\Rightarrow\ K_0 \text{ collapses}.
\]

% ---------------------------------------------------------------
\subsubsection{P3: Predictions Concerning Transitions and Dimensionality}

\paragraph{(P3.1) No Dimensional Self-Generation.}
OC predicts:
\[
K_0 \centernot\Rightarrow K_1
\]
unless the meta-space $M_0$ provides an admissible axis $A_1\in M_0$.  
Thus spontaneous dimensionality increase is forbidden.

\paragraph{(P3.2) Transition Requires an Axis Seed.}
A transition  
\[
K_0 \longrightarrow K_1
\]
is possible only if an operator $\Psi_{0\to1}$ is defined on $M_0$,
  
mapping $(S,\Delta,\mathcal{C})$ to a field $\phi(x)$ over a newly born axis.  
Therefore:
\[
\neg\exists A_1 \in \Omega(M_0)
\quad\Rightarrow\quad
K_0 \not\to K_1.
\]

\paragraph{(P3.3) No Temporal Structure Prediction.}
Since $K_0$ contains no time axis,
\[
\tau \notin A(K_0).
\]
Thus $K_0$ predicts that no temporal evolution is definable at this level.  
All dynamics are emergent properties of $K_1$.

% ---------------------------------------------------------------
\subsubsection{P4: Predictions Concerning Tension and Collapse}

\paragraph{(P4.1) Zero Tension is Necessary.}
Since $T(K_0)$ is undefined (no potentials, no axes),  
the only admissible value is structural quasi-zero:
\[
T_0 = 0.
\]
Thus $K_0$ predicts that no internal tension may accumulate.

\paragraph{(P4.2) Collapse Trigger Prediction.}
If compositional inconsistencies accumulate (violation of C1–C4′),
\[
\exists x,y,z: \mathcal{C}(x,y), \mathcal{C}(y,z), \text{ but } \mathcal{C}(x,z) \text{ undefined}
\]
and this violates minimal coherence constraints, then:
\[
K_0 \to \varnothing.
\]

\paragraph{(P4.3) No Partial Collapse.}
Predicted:
\[
K_0 \text{ exists} \Rightarrow \text{fully coherent};
\]
there is no notion of partial existence because $k$ is undefined at this level.

% ---------------------------------------------------------------
\subsubsection{P5: Predictions Concerning Compatibility with Meta-Spaces}

\paragraph{(P5.1) Meta-space Incompleteness Prediction.}
OC predicts:
\[
\Omega(M_0) \subsetneq \Omega(M_1)
\]
by the Theorem of Incomplete Meta-spaces.  
Thus $M_0$ cannot be maximal.

\paragraph{(P5.2) Forbidden Full Closure.}
$M_0$ cannot contain the full structure of $K_1$:
\[
A_1 \notin M_0,\quad \Delta_1 \notin M_0.
\]

\paragraph{(P5.3) Boundary Prediction.}
Since boundaries are undefined in $K_0$,
\[
\partial\Omega(K_0) = \varnothing,
\]
predicting that boundaries only emerge at $K_1$ with the birth of the first axis.

% ---------------------------------------------------------------
\subsubsection{Summary}

The predictions for $K_0$ state that:

\begin{itemize}
    \item proto-continua must possess nonzero structural slack $\Theta_0>0$;
    \item compositional structure must be partial and cannot be globally defined;
    \item no dimensionality, no time and no dynamics can emerge from $K_0$ internally;
    \item transitions to $K_1$ require admissibility from $M_0$;
    \item collapse occurs whenever minimal coherence fails;
    \item boundaries and axes are impossible at this level.
\end{itemize}

These predictions provide the invariant conditions under which  
the emergence, coherence and eventual transition of proto-continua  
are permitted within the OC framework.
