% OC Core 1.1 — Extended modules and cross-level structures

\section*{Extended modules and cross-level structures}
\addcontentsline{toc}{section}{Extended modules and cross-level structures}
\label{sec:modules_master}

This section provides a high-level overview of the extended structural modules
of the Ontology of Continua. Each module is developed in its own section of
the Core; here we only summarise their roles and relations.

The extended modules are:

- \textbf{K-levels master} (Sec.~\ref{sec:klevels-master}): 
  global overview of all continua \(K_0\)–\(K_{12}\), their state spaces,
  axes, thresholds and core processes.

- \textbf{Embedding spaces / M-spaces} (Sec.~\ref{sec:mspaces-overview}):
  definition of meta-spaces \(M_x\) that constrain and parameterise the
  evolution of continua \(K_x\).

- \textbf{Cross-level / cross-K structures} (Sec.~\ref{sec:crossk_master}):
  operators, mappings and landscape structures that link different levels
  of the hierarchy and generate phase transitions between them.

- \textbf{Processes} (Sec.~\ref{sec:processes-master}):
  unified description of evolution operators \(E = (F,G,H,Q,R,S,U)\) across
  levels, including tension dynamics, thresholds, flows and cycles.

- \textbf{Cycles} (Sec.~\ref{sec:cycles-master}):
  structural and dynamical cycles \(C_j\) as carriers of continuity,
  memory and temporal organisation.

- \textbf{Predictions} (Sec.~\ref{sec:predictions-master}):
  level-specific and cross-level predictions, empirical signatures and
  qualitative regimes implied by the formalism.

- \textbf{Experiments} (Sec.~\ref{sec:experiments-master}):
  conceptual and concrete experimental designs to probe continua,
  transitions and threshold landscapes.

- \textbf{Falsifiability structure} (Sec.~\ref{sec:falsifiability-master}):
  explicit conditions under which the Ontology of Continua can be
  refuted, constrained or forced to restructure its core.

- \textbf{Jets / local evolution structure} (Sec.~\ref{sec:jets-master}):
  local expansions of the evolution operator and short-time dynamics
  around given configurations of \(K_x\) and \(M_x\).

Together, these modules refine and operationalise the core formalism across
levels and applications, while maintaining a single coherent hierarchy of
continua and meta-spaces.
