% ================================================================
% ==== FILE: content/m_spaces/m11.tex
% ================================================================

% ==============================
%  Ontology of Continua — Core
%  Meta-Space: M11
%  FULL MODULE — FINAL VERSION
% ==============================

\subsubsection{$M_{11}$ Overview}
\label{sec:m11-overview}

$M_{11}$ is the highest meta-space required for continua up to level $K_{11}$.  
Where $M_{10}$ hosts categories of model-categories (meta-theoretical structures),  
$M_{11}$ hosts \emph{hierarchies of meta-theoretical towers} and provides  
the global logical environment for their compatibility, reflection,  
and higher-order stabilisation.

In this space, the fundamental units are \emph{meta-meta-categorical} objects:
structured towers
\[
(\mathcal{C}^{(0)}, \mathcal{C}^{(1)}, \ldots, \mathcal{C}^{(n)}, \ldots)
\]
where each $\mathcal{C}^{(i)}$ is itself a category-of-categories at level $i$
and morphisms between towers are transformations that preserve their full  
vertical structure.

Thus $M_{11}$ is the minimal logical environment in which  
\emph{meta-hierarchical continua} can exist.

% ---------------------------------------------------------------
\subsubsection*{1. Admissible Configuration Space $\Omega(M_{11})$}

\[
\Omega(M_{11}) =
\left\{
\left( \mathbf{T},\ \mathfrak{F},\ A_{11},\
\Theta_{11},\ C_{11},\Sigma_{11} \right)
\ \middle|\ 
\text{global hierarchy-coherence holds}
\right\}.
\]

Where:

\begin{itemize}
    \item $\mathbf{T}$ — admissible infinite or finite towers of meta-categories,
    \item $\mathfrak{F}$ — meta-functors acting on towers,
    \item $A_{11}$ — axes of meta-hierarchical variation,
    \item $\Theta_{11}$ — global coherence thresholds at the highest level,
    \item $C_{11}$ — cycles of trans-hierarchical self-organisation,
    \item $\Sigma_{11}$ — structural constraints defining $M_{11}$.
\end{itemize}

Admissibility requires:

\begin{enumerate}
    \item hierarchy must close under vertical composition,
    \item no divergence of coherence across levels,
    \item existence of at least one stable trans-hierarchical cycle,
    \item compatibility with $M_{10}$ through canonical projection 
          $\pi_{10}:M_{11}\to M_{10}$,
    \item finiteness of global meta-tension.
\end{enumerate}

\[
\Omega(K_{11}) \subseteq \Omega(M_{11}) 
\quad\text{(Theorem 8: necessity and sufficiency of compatibility)}.
\]

% ---------------------------------------------------------------
\subsubsection*{2. Axes $A(M_{11})$}

\[
A(M_{11}) =
\{
A_{\mathrm{tower}},\
A_{\mathrm{coherence}},\
A_{\mathrm{depth}},\
A_{\mathrm{reflection}},\
A_{\mathrm{translation}},\
A_{\mathrm{hierarchical\ consistency}},\
A_{\mathrm{global}}
\}.
\]

Interpretation:

\begin{itemize}
    \item $A_{\mathrm{tower}}$ — variation across levels of the meta-theoretical tower,
    \item $A_{\mathrm{coherence}}$ — maintenance of coherence across multiple levels,
    \item $A_{\mathrm{depth}}$ — hierarchical depth and structural recursion,
    \item $A_{\mathrm{reflection}}$ — self-description across levels,
    \item $A_{\mathrm{translation}}$ — translation between different meta-hierarchies,
    \item $A_{\mathrm{hierarchical\ consistency}}$ — preservation of structural rules across the tower,
    \item $A_{\mathrm{global}}$ — axes governing global admissibility of entire hierarchies.
\end{itemize}

Key feature:

\[
\dim(M_{11}) > \dim(M_{10})
\quad\text{(Theorem: monotonicity and irreversibility of dimension growth)}.
\]

% ---------------------------------------------------------------
\subsubsection*{3. Potentials $P(M_{11})$}

\[
P(M_{11}) =
\{
F_{\mathrm{global\ coherence}},\
F_{\mathrm{hierarchical\ depth}},\
F_{\mathrm{meta\mbox{-}reflection}},\
F_{\mathrm{compatibility}},\
F_{\mathrm{translation}},\
F_{\mathrm{unification}},\
F_{\mathrm{stability}}
\}.
\]

Interpretation:

\begin{itemize}
    \item $F_{\mathrm{global\ coherence}}$ — ability to maintain 
        coherence across all levels,
    \item $F_{\mathrm{hierarchical\ depth}}$ — capacity to sustain deep tower structures,
    \item $F_{\mathrm{meta\mbox{-}reflection}}$ — ability to reflect on entire hierarchies,
    \item $F_{\mathrm{compatibility}}$ — harmony between levels,
    \item $F_{\mathrm{translation}}$ — mapping between different towers,
    \item $F_{\mathrm{unification}}$ — unification across hierarchical scales,
    \item $F_{\mathrm{stability}}$ — overall stability of meta-hierarchical dynamics.
\end{itemize}

Existence of $K_{11}$ requires:

\[
F_{\mathrm{global\ coherence}} > \Theta_{11}^{\mathrm{collapse}},
\qquad
F_{\mathrm{meta\mbox{-}reflection}} > \Theta_{11}^{\mathrm{self\mbox{-}break}}.
\]

% ---------------------------------------------------------------
\subsubsection*{4. Thresholds $\Theta(M_{11})$}

Critical thresholds:

\begin{itemize}
    \item $\Theta_{11}^{\mathrm{coherence}}$ — minimal global coherence,
    \item $\Theta_{11}^{\mathrm{depth}}$ — minimal structural depth,
    \item $\Theta_{11}^{\mathrm{translation}}$ — threshold for interoperability,
    \item $\Theta_{11}^{\mathrm{collapse}}$ — boundary where the hierarchy fails,
    \item $\Theta_{11}^{\mathrm{self\mbox{-}break}}$ — self-reference instability.
\end{itemize}

Crossing $\Theta_{11}^{\mathrm{collapse}}$ yields hierarchical fragmentation:

\[
\mathbf{T} \to (\mathcal{C}^{(i)}) \quad \text{without admissible vertical structure}.
\]

% ---------------------------------------------------------------
\subsubsection*{5. Flows $J(M_{11})$}

\[
J(M_{11}) =
(
J_{\mathrm{tower}},\
J_{\mathrm{coherence}},\
J_{\mathrm{translation}},\
J_{\mathrm{meta\mbox{-}inference}},\
\nabla_i T_{11},\
J_{\mathrm{reflection}},\
J_{\mathrm{global}}
).
\]

Interpretation:

\begin{itemize}
    \item $J_{\mathrm{tower}}$ — flows acting on hierarchical depth,
    \item $J_{\mathrm{coherence}}$ — flows restoring multi-level consistency,
    \item $J_{\mathrm{translation}}$ — flows mapping one tower into another,
    \item $J_{\mathrm{meta\mbox{-}inference}}$ — inference acting across the full tower,
    \item $\nabla_i T_{11}$ — gradients of global meta-tension,
    \item $J_{\mathrm{reflection}}$ — recursion and reflection across levels,
    \item $J_{\mathrm{global}}$ — flows determining global admissibility.
\end{itemize}

% ---------------------------------------------------------------
\subsubsection*{6. Cycles $C(M_{11})$}

Fundamental cycles:

\begin{itemize}
    \item $C_{\mathrm{hierarchy}}$ — formation, refinement and stabilisation of the tower,
    \item $C_{\mathrm{coherence}}$ — coherence restoration loops across levels,
    \item $C_{\mathrm{translation}}$ — mapping towers and verifying equivalence,
    \item $C_{\mathrm{reflection}}$ — recursive cycles enabling 
          understanding of the hierarchy from within,
    \item $C_{\mathrm{global}}$ — global admissibility cycles ensuring
          $\Omega(K_{11})\subseteq\Omega(M_{11})$.
\end{itemize}

Criterion of life:

\[
\oint_{C_{\mathrm{global}}} dA_{\mathrm{global}} \approx 0.
\]

% ---------------------------------------------------------------
\subsubsection*{7. Boundary $\partial\Omega(M_{11})$}

Near the boundary:

\begin{itemize}
    \item hierarchical divergence,
    \item loss of coherence between levels,
    \item unresolvable contradictions,
    \item runaway recursion,
    \item breakdown of global admissibility.
\end{itemize}

Crossing $\partial\Omega(M_{11})$ forces collapse into $M_{10}$-admissible structures.

% ---------------------------------------------------------------
\subsubsection*{8. Relation to $M_{10}$ and higher meta-spaces}

\textbf{Relation to $M_{10}$:}

\[
A_{\mathrm{hierarchical\ consistency}} \in A(M_{11}) \setminus A(M_{10}),
\]
thus $M_{11}$ strictly extends $M_{10}$.

Vertical projection:

\[
\pi_{10}: M_{11} \to M_{10}
\]
forgets hierarchical depth and retains only meta-theoretical structure.

\textbf{Towards $M_{12}$ (if defined):}

$M_{12}$ would host universal meta-meta-logical structures and  
global constraints on entire families of meta-hierarchies.

$M_{11}$ is the minimal space where such extension becomes admissible.

