% ================================================================
% ==== FILE: content/m_spaces/m12.tex
% ================================================================

% ==============================
%  Ontology of Continua — Core
%  Meta-Space: M12
%  FULL MODULE — FINAL VERSION
% ==============================

\subsubsection{$M_{12}$ Overview}
\label{sec:m12-overview}

$M_{12}$ is the highest meta-space required for continua up to level $K_{12}$.  
While $M_{11}$ hosts structured towers of meta-theoretical hierarchies,  
$M_{12}$ hosts \emph{universal meta-logical environments} governing the  
admissibility, coherence and global constraints for entire \emph{families of  
meta-hierarchies}.

In short:  
\[
M_{12} = \text{the minimal meta-logical space in which the entire tower }
M_1, \ldots, M_{11} \text{ can coexist coherently.}
\]

It is not simply “one level higher”;  
it introduces a new class of differences (Ax. 21),  
a new axis unavailable to $M_{11}$,  
and a global constraint governing all previous M-spaces.

This makes $M_{12}$ the terminal environment for Core~1.1.

% ---------------------------------------------------------------
\subsubsection*{1. Admissible Configuration Space $\Omega(M_{12})$}

\[
\Omega(M_{12}) =
\left\{
\left( \mathfrak{U}, A_{12}, \Theta_{12},
C_{12}, \Lambda_{12}, T_{12} \right)
\ \middle|\
\text{global meta-logical admissibility holds}
\right\}.
\]

Where:

\begin{itemize}
    \item $\mathfrak{U}$ — admissible \emph{universes of meta-hierarchies},  
          i.e. collections of towers from $M_{11}$ equipped with compatibility relations;
    \item $A_{12}$ — axes determining variation across universes of meta-hierarchies;
    \item $\Theta_{12}$ — global thresholds ensuring logical admissibility;
    \item $C_{12}$ — cycles governing stabilisation at the universal level;
    \item $\Lambda_{12}$ — universal constraints (logical, structural, categorical);
    \item $T_{12}$ — global structural tension across families of hierarchies.
\end{itemize}

Admissibility requires:

\begin{enumerate}
    \item coherence of all embedded meta-hierarchies (\(M_1\)–\(M_{11}\)),
    \item existence of at least one universal stabilising cycle,
    \item global consistency with the operator $\Psi$ (meta-space generation),
    \item bounded meta-tension $T_{12}<\infty$,
    \item preservation of universal structural laws $\Lambda_{12}$.
\end{enumerate}

By Theorem 8:

\[
\Omega(K_{12}) \subseteq \Omega(M_{12})
\quad \text{is necessary and sufficient for the existence of }K_{12}.
\]

% ---------------------------------------------------------------
\subsubsection*{2. Axes $A(M_{12})$}

\[
A(M_{12}) =
\{
A_{\mathrm{universal}},\
A_{\mathrm{meta\mbox{-}logical}},\
A_{\mathrm{cross\mbox{-}hierarchy}},\
A_{\mathrm{invariance}},\
A_{\mathrm{global\ compatibility}},\
A_{\mathrm{admissibility}}
\}.
\]

Where:

\begin{itemize}
    \item $A_{\mathrm{universal}}$ — variation across entire universes of meta-hierarchies,
    \item $A_{\mathrm{meta\mbox{-}logical}}$ — global logical constraints,
    \item $A_{\mathrm{cross\mbox{-}hierarchy}}$ — variation across $M_i$ structures,
    \item $A_{\mathrm{invariance}}$ — invariants preserved across all levels,
    \item $A_{\mathrm{global\ compatibility}}$ — compatibility axis for full K/M stacks,
    \item $A_{\mathrm{admissibility}}$ — the existence axis for $K_{12}$ itself.
\end{itemize}

This introduces a new class of differences not embeddable into $M_{11}$:

\[
A_{\mathrm{universal}}\notin A(M_{11}),
\]
thus by Ax. 21:

\[
\dim(M_{12}) > \dim(M_{11}).
\]

% ---------------------------------------------------------------
\subsubsection*{3. Potentials $P(M_{12})$}

\[
P(M_{12}) =
\{
U_{\mathrm{universal}},\
U_{\mathrm{coherence}},\
U_{\mathrm{global\ inference}},\
U_{\mathrm{meta\mbox{-}stability}},\
U_{\mathrm{admissibility}},\
U_{\mathrm{invariant\ preservation}}
\}.
\]

Interpretation:

\begin{itemize}
    \item $U_{\mathrm{universal}}$ — potential supporting whole universes of meta-hierarchies,
    \item $U_{\mathrm{coherence}}$ — ability to maintain coherence across all embedded M-spaces,
    \item $U_{\mathrm{global\ inference}}$ — potential for reasoning across universes,
    \item $U_{\mathrm{meta\mbox{-}stability}}$ — ability to resist collapse of entire towers,
    \item $U_{\mathrm{admissibility}}$ — ensures $\Omega(K_{12})\subseteq\Omega(M_{12})$,
    \item $U_{\mathrm{invariant\ preservation}}$ — ensures invariants across levels.
\end{itemize}

$K_{12}$ exists iff:

\[
U_{\mathrm{admissibility}} > \Theta_{12}^{\mathrm{existence}}.
\]

% ---------------------------------------------------------------
\subsubsection*{4. Thresholds $\Theta(M_{12})$}

Critical thresholds:

\[
\Theta_{12}=
\{
\Theta_{12}^{\mathrm{coherence}},\
\Theta_{12}^{\mathrm{universal}},\
\Theta_{12}^{\mathrm{collapse}},\
\Theta_{12}^{\mathrm{invariance}},\
\Theta_{12}^{\mathrm{reflection}}
\}.
\]

Interpretation:

\begin{itemize}
    \item $\Theta_{12}^{\mathrm{coherence}}$ — minimal coherence among all $M_i$,
    \item $\Theta_{12}^{\mathrm{universal}}$ — minimal universal compatibility,
    \item $\Theta_{12}^{\mathrm{collapse}}$ — threshold where entire meta-hierarchies disintegrate,
    \item $\Theta_{12}^{\mathrm{invariance}}$ — minimal preservation of invariants,
    \item $\Theta_{12}^{\mathrm{reflection}}$ — threshold for self-consistency of meta-logic.
\end{itemize}

Crossing $\Theta_{12}^{\mathrm{collapse}}$ yields:

\[
\mathfrak{U} \to \{ M_i \}_{i=1}^{11}
\quad\text{with loss of universal structure}.
\]

% ---------------------------------------------------------------
\subsubsection*{5. Flows $J(M_{12})$}

\[
J(M_{12})=
(
J_{\mathrm{universal}},\
J_{\mathrm{coherence}},\
J_{\mathrm{reflection}},\
J_{\mathrm{meta\mbox{-}inference}},\
J_{\mathrm{cross\mbox{-}hierarchy}},\
\nabla T_{12},\
J_{\mathrm{stability}}
).
\]

Interpretation:

\begin{itemize}
    \item $J_{\mathrm{universal}}$ — flows reorganising universes of meta-hierarchies,
    \item $J_{\mathrm{coherence}}$ — flows restoring coherence across $M_1$–$M_{11}$,
    \item $J_{\mathrm{reflection}}$ — self-referential meta-logical flows,
    \item $J_{\mathrm{meta\mbox{-}inference}}$ — inference acting on entire universes,
    \item $J_{\mathrm{cross\mbox{-}hierarchy}}$ — stabilisation across different M-levels,
    \item $\nabla T_{12}$ — gradients of global structural tension,
    \item $J_{\mathrm{stability}}$ — flows preventing collapse of universal structure.
\end{itemize}

% ---------------------------------------------------------------
\subsubsection*{6. Cycles $C(M_{12})$}

Fundamental cycles:

\begin{itemize}
    \item $C_{\mathrm{universal}}$ — generation and stabilisation of universes of meta-hierarchies,
    \item $C_{\mathrm{coherence}}$ — coherence-restoring loops across all M-spaces,
    \item $C_{\mathrm{reflection}}$ — recursive cycles maintaining meta-logical integrity,
    \item $C_{\mathrm{invariance}}$ — cycles ensuring global invariants,
    \item $C_{\mathrm{admissibility}}$ — cycles ensuring existence of $K_{12}$.
\end{itemize}

Life criterion:

\[
\oint_{C_{\mathrm{universal}}} dA_{\mathrm{universal}} \approx 0.
\]

% ---------------------------------------------------------------
\subsubsection*{7. Boundary $\partial\Omega(M_{12})$}

Near the boundary:

\begin{itemize}
    \item universal incompatibility,
    \item breakdown of global structural invariants,
    \item meta-logical incoherence,
    \item collapse of stabilising cycles,
    \item divergence of meta-tension.
\end{itemize}

Crossing the boundary forces reduction to isolated $M_{11}$ universes.

% ---------------------------------------------------------------
\subsubsection*{8. Relation to $M_{11}$ and full K/M hierarchy}

\[
M_{11} \subset M_{12}
\quad\text{strictly}.
\]

Projection:

\[
\pi_{11}: M_{12} \to M_{11}
\]
forgets universal constraints and retains only tower-structured hierarchies.

$M_{12}$ is the minimal environment in which:

\[
\Psi: M_{11} \to M_{12}
\]
is well-defined (Run 11).

The existence of $K_{12}$ requires:

\[
\Omega(K_{12}) \subseteq \Omega(M_{12}).
\]

$M_{12}$ is thus the terminal meta-space for the Core.

