% ======================================================================
% Ontology of Continua — Core 1.1
% Cross-K module: crossk_k7_k8.tex
% Cross-level relations between K7 and K8
% ======================================================================

\subsubsection{K7–K8 cross-level structure}
\label{sec:crossk-k7-k8}

The transition $K_7 \rightarrow K_8$ marks the emergence of the
civilisational–technological continuum from the social continuum.
Where $K_7$ provides norms, roles, communication flows and institutional
cycles, $K_8$ introduces:
\begin{itemize}
  \item technological and infrastructural axes,
  \item large-scale symbolic and informational systems,
  \item energy and logistics networks,
  \item economic mechanisms of reproduction,
  \item repair, maintenance and high-order cooperation cycles,
  \item global continuity conditions for population-level systems.
\end{itemize}

$K_8$ is the first continuum capable of sustaining persistent structures
that exceed the cognitive and social horizons of individual agents.

% ----------------------------------------------------------------------
\subsubsection{1. Levels involved and their roles}

\paragraph{K7 (social continuum).}
$K_7$ provides:
\begin{itemize}
  \item social norms and roles ($A_{\mathrm{soc}}$),
  \item communication flows $J_{\mathrm{comm}}$,
  \item institutional cycles $C_{\mathrm{inst}}$,
  \item collective potentials $P_{\mathrm{soc}}$ (trust, legitimacy),
  \item cohesion and enforcement thresholds $\Theta_{\mathrm{soc}}$.
\end{itemize}

However, $K_7$ lacks:
\begin{itemize}
  \item large-scale infrastructures,
  \item technological production cycles,
  \item energy-density thresholds,
  \item material and informational logistics,
  \item stable multi-generational reproduction mechanisms.
\end{itemize}

\paragraph{K8 (civilisational–technological continuum).}
$K_8$ introduces:
\begin{itemize}
  \item axes for infrastructure, technology and symbolic systems ($A_{\mathrm{tech}}$),
  \item potentials for energy density, repair capacity, logistics throughput,
  \item flows $J^{(8)} = (J_{\mathrm{energy}}, J_{\mathrm{material}},
        J_{\mathrm{logistics}}, J_{\mathrm{info}}, J_{\mathrm{population}})$,
  \item infrastructure cycles $C_{\mathrm{sys}}$ (production → transport → use → maintenance),
  \item thresholds $\Theta_8 = (\Theta_E, \Theta_L, \Theta_I, \Theta_M, \Theta_R)$,
  \item measure $k_8$ describing civilisational continuity.
\end{itemize}

Thus $K_8$ is the first continuum with large-scale energetic and informational
autonomy.

% ----------------------------------------------------------------------
\subsubsection{2. Shared and inherited axes and thresholds}

\paragraph{Shared symbolic structures.}
Symbolic structures and communication systems of $K_7$ become the basis for:
\[
A_{\mathrm{symbol}}(K_7) \rightarrow A_{\mathrm{tech}}(K_8),
\]
enabling technical codification (writing, mathematics, engineering,
standards).

\paragraph{New axes in \texorpdfstring{$K_8$}{K_8}.}
These include:
\begin{itemize}
  \item $A_{\mathrm{infra}}$ — roads, power grids, water systems,
  \item $A_{\mathrm{energy}}$ — energy density and conversion systems,
  \item $A_{\mathrm{econ}}$ — economic mechanisms of reproduction,
  \item $A_{\mathrm{info}}$ — information storage and transmission,
  \item $A_{\mathrm{repair}}$ — maintenance and regeneration structures.
\end{itemize}

\paragraph{Thresholds.}
Inherited social cohesion thresholds remain necessary but insufficient.
New thresholds arise:
\begin{itemize}
  \item $\Theta_E$ — minimal energy density for system-level reproduction,
  \item $\Theta_L$ — logistics throughput thresholds,
  \item $\Theta_I$ — informational coherence thresholds,
  \item $\Theta_M$ — repair capacity threshold,
  \item $\Theta_R$ — demographic sustainability threshold.
\end{itemize}

If any of these thresholds fail, $\Omega(K_8)$ collapses.

% ----------------------------------------------------------------------
\subsubsection{3. Cross-level flows, cycles, and tensions}

\paragraph{Flows.}
Flows in $K_8$ generalise and scale-up social flows:
\[
J^{(8)} = (J_{\mathrm{energy}}, J_{\mathrm{material}}, 
          J_{\mathrm{logistics}}, J_{\mathrm{info}}, 
          J_{\mathrm{population}}).
\]

\paragraph{Cycles.}
Civilisational stability requires the infrastructure cycle $C_{\mathrm{sys}}$:
\[
\text{production} \rightarrow \text{transport} \rightarrow
\text{use} \rightarrow \text{maintenance} \rightarrow \text{production}.
\]
Stability condition:
\[
\oint dR \approx 0.
\]

\paragraph{Tension.}
Cross-level tension:
\[
T_{7,8} =
f(\Theta_E - P_{\mathrm{energy}},\ 
  \Theta_L - J_{\mathrm{logistics}},\ 
  \Theta_I - P_{\mathrm{info}},\ 
  \Theta_M - P_{\mathrm{repair}},\
  \Theta_R - P_{\mathrm{population}}).
\]

Excess tension destroys infrastructural continuity.

% ----------------------------------------------------------------------
\subsubsection{4. Birth and death conditions across the levels}

\paragraph{Birth of \texorpdfstring{$K_8$}{K_8}.}
$K_8$ emerges when:
\[
T_7 > \Theta_{7,\mathrm{dim}}
\quad\text{and}\quad
A_{\mathrm{tech}}, A_{\mathrm{infra}} \in M_8 \setminus A(K_7).
\]

The domain expands:
\[
\Omega(K_8) = \Omega(K_7) \cup \Delta\Omega_{\mathrm{civ}}.
\]

\paragraph{Death propagation.}
If social thresholds fail ($\Theta_{\mathrm{soc}}$), civiliational systems lose
coherence and collapse to $K_7$.

If any of the five $K_8$ thresholds fail:
\[
\Theta_E,\Theta_L,\Theta_I,\Theta_M,\Theta_R,
\]
then $\Omega(K_8)=\varnothing$.

$K_7$ may remain stable unless its own thresholds are violated.

% ----------------------------------------------------------------------
\subsubsection{5. Conceptual examples}

\paragraph{Energy-density escalation.}
When energy density surpasses $\Theta_E$, technological and economic
structures become self-sustaining.

\paragraph{Information codification.}
Social symbols evolve into technical systems (writing → mathematics →
science), enabling new axes in $A_{\mathrm{tech}}$.

\paragraph{Infrastructure failure.}
If logistics capacity drops below $\Theta_L$ or repair rates below $\Theta_M$,
civilisational cycles break and the system collapses:
\[
C_{\mathrm{sys}} \nrightarrow \text{closed loop}.
\]

Thus the K7→K8 transition can be summarised as:
\[
\textbf{The emergence of technological–civilisational structure from social
norms, communication flows and institutional stability.}
\]

