% ======================================================================
% Ontology of Continua — Core 1.1
% Cross-K module: crossk_k10_k11.tex
% Cross-level relations between K10 and K11
% ======================================================================

\subsubsection{K10–K11 cross-level structure}
\label{sec:crossk-k10-k11}

The transition $K_{10} \rightarrow K_{11}$ marks the emergence of a
\textbf{trans–metatheoretical continuum} from the metatheoretical
continuum. While $K_{10}$ provides functorial mappings, meta-logic and
categorical structures that relate theoretical systems, $K_{11}$ introduces:
\begin{itemize}
  \item universal operator classes acting across multiple meta-frameworks,
  \item higher-order compatibility conditions for meta-models,
  \item global thresholds for trans-meta coherence,
  \item flows that restructure entire families of meta-models,
  \item cycles that regulate meta-level evolution itself,
  \item a constraint layer ensuring that model evolution across all levels
        $K_0\dots K_{10}$ remains consistent within a higher envelope.
\end{itemize}

In Core~1.1, $K_{11}$ is not fully formalised; this section provides only its
structural role within the cross-level architecture.

% ----------------------------------------------------------------------
\subsubsection{1. Levels involved and their roles}

\paragraph{K10 (metatheoretical continuum).}
$K_{10}$ provides:
\begin{itemize}
  \item axes $A^{10}$ (functors, meta-logic, compatibility categories),
  \item potentials $P^{10}$ (meta-coherence, expressivity, universality),
  \item thresholds $\Theta^{10}$ (diagrammatic consistency,
        meta-logical well-formedness),
  \item flows $J^{10}$ (interpretation, translation, reconstruction),
  \item cycles $C^{10}$ (meta-stability, refinement loops).
\end{itemize}

\paragraph{K11 (trans–metatheoretical continuum).}
$K_{11}$, in its minimal Core~1.1 formulation, introduces:
\begin{itemize}
  \item axes $A^{11}$ describing relations between entire classes of
        meta-frameworks,
  \item potentials $P^{11}$ quantifying global coherence across meta-levels,
  \item thresholds $\Theta^{11}$ for trans-meta consistency and
        compatibility,
  \item flows $J^{11}$ restructuring families of metamodels,
  \item cycles $C^{11}$ regulating the evolution of $K_{10}$ structures,
  \item a measure $k(K_{11})$ for universal structural coherence.
\end{itemize}

$K_{11}$ serves as a constraint layer for all lower modelling levels.

% ----------------------------------------------------------------------
\subsubsection{2. Shared / inherited axes and thresholds}

\paragraph{Inherited structure.}
Metamodels in $K_{10}$ become objects; $K_{11}$ introduces morphisms between
meta-morphisms:
\[
\mathsf{MetaFunctors}(K_{10}) \rightarrow \mathsf{TransFunctors}(K_{11}).
\]

\paragraph{New axes.}
Axes in $A^{11}$ include:
\begin{itemize}
  \item $A_{\mathrm{ultra}}$ — operator classes acting on categories of
        meta-models,
  \item $A_{\mathrm{compat}}^{(11)}$ — trans-meta compatibility relations,
  \item $A_{\mathrm{reflect}}^{(11)}$ — higher-order self-reflection axes,
  \item $A_{\mathrm{limit}}$ — constraints from universal consistency.
\end{itemize}

\paragraph{New thresholds.}
$\Theta^{11}$ includes:
\begin{itemize}
  \item \textbf{trans-meta coherence threshold} — failure collapses $K_{11}$,
  \item \textbf{compatibility threshold} — incompatible meta-frameworks
        cannot coexist,
  \item \textbf{universal consistency threshold} — prohibits paradox at the
        trans-meta level,
  \item \textbf{expressivity threshold} — minimal representational sufficiency
        across $K_0\dots K_{10}$.
\end{itemize}

Violation of $\Theta^{10}$ prevents the emergence of $K_{11}$.

% ----------------------------------------------------------------------
\subsubsection{3. Cross-level flows, cycles, and tensions}

\paragraph{Flows.}
Flows in $K_{11}$ extend meta-flows:
\[
J^{11} = (J_{\mathrm{ultra}},\ J_{\mathrm{crossmeta}},\
         J_{\mathrm{lift2}},\ J_{\mathrm{constraint}}),
\]
where:
\begin{itemize}
  \item $J_{\mathrm{ultra}}$ — restructuring of categories of metamodels,
  \item $J_{\mathrm{crossmeta}}$ — translation between incompatible meta-levels,
  \item $J_{\mathrm{lift2}}$ — lifting meta-frameworks into $K_{11}$,
  \item $J_{\mathrm{constraint}}$ — enforcing global structural limits.
\end{itemize}

\paragraph{Cycles.}
$C^{11}$ includes higher-order cycles such as:
\[
C_{\mathrm{ultra}}:\ 
\text{metamodel} \rightarrow \text{lift} \rightarrow \text{constraint}
\rightarrow \text{refinement} \rightarrow \text{metamodel}.
\]

These cycles regulate evolution of the meta-continuum.

\paragraph{Tension.}
Cross-level tension:
\[
T_{10,11} =
f(\Theta^{11} - P^{11},\ A^{11} - A^{10},\
  J^{11} - J^{10},\ \Theta^{10} - \Theta^{11}).
\]

Excess tension collapses $K_{11}$ to a simpler metatheoretical regime.

% ----------------------------------------------------------------------
\subsubsection{4. Birth and death conditions across the levels}

\paragraph{Birth of $K_{11}$.}
$K_{11}$ emerges when:
\[
T_{10} > \Theta_{10,\mathrm{dim}}
\quad\text{and}\quad
A^{11} \in M_{11} \setminus A(K_{10}),
\]
i.e. when distinctions between entire meta-framework classes become necessary.

Expansion:
\[
\Omega(K_{11}) = \Omega(K_{10}) \cup \Delta\Omega_{\mathrm{ultra}}.
\]

\paragraph{Death propagation.}
If $\Theta^{10}$ fails, the trans-meta layer cannot exist.

If $\Theta^{11}$ fails (incompatibility, meta-paradox), $K_{11}$ collapses but
$K_{10}$ persists:
\[
\Omega(K_{11}) \rightarrow \varnothing.
\]

% ----------------------------------------------------------------------
\subsubsection{5. Conceptual examples}

\paragraph{Reconciling incompatible meta-frameworks.}
$K_{11}$ can provide a higher-level compatibility operator that relates
meta-frameworks which cannot directly map into one another in $K_{10}$.

\paragraph{Universal constraints.}
$K_{11}$ enforces global consistency across all modelling levels:
\[
K_{11} : \{K_0,\dots,K_{10}\} \rightarrow \text{coherent envelope}.
\]

\paragraph{Collapse case.}
If a trans-meta inconsistency arises:
\[
T_{10,11} > \Theta^{11}_{\mathrm{coh}},
\]
$K_{11}$ collapses into $K_{10}$.

Thus the K10→K11 transition can be summarised as:
\[
\textbf{The emergence of universal operators and constraints capable
of relating and regulating entire classes of metatheoretical frameworks.}
\]

