% ================================================================
% ==== FILE: content/falsifiability/falsifiability_k2.tex
% ================================================================

\section{Falsifiability of \texorpdfstring{$K_2$}{K_2}}
\label{sec:falsifiability-k2}

The continuum $K_2$ is the first fully physical continuum in the OC
hierarchy.  
It is characterised by:
\begin{itemize}
    \item the emergence of a new axis $A_{\mathrm{phys}}$ not contained
    in $\mathrm{span}(A_1)$,
    \item the birth of physical configuration space $\Omega(K_2)$ defined
    by connectivity, percolation and local interaction,
    \item the existence of a critical threshold $p_c$ determining global
    connectedness,
    \item the appearance of physical potentials and flows
    ($P_{\mathrm{phys}}, J_{\mathrm{phys}}$),
    \item physically meaningful boundaries $\partial\Omega(K_2)$ defined
    by field constraints,
    \item non-trivial structural tension $T_2$.
\end{itemize}

Falsifiability of $K_2$ concerns the possibility that any of these
structures fail.



% ================================================================
\subsection{Topological and Connectivity Falsifiability}
% ================================================================

$K_2$ crucially depends on global connectivity.  
The physical continuum is defined by a connected region emerging after a
percolation-like transition.

It is falsified if:

\paragraph{(F2.1) No global connected component exists.}
If the system remains below the percolation threshold $p_c$ for all
configurations, then
\[
\mathrm{Conn}(\Omega(K_2)) = 0
\]
and the continuum cannot exist.

\paragraph{(F2.2) Percolation theory predicts no admissible connected
phase.}
If the structure of $K_1$ or the geometry of admissible states forbids a
transition through $p_c$, $K_2$ cannot arise.

\paragraph{(F2.3) Connectivity changes violate stability.}
If connected components appear and disappear without a stable regime,
$K_2$ becomes dynamically inconsistent.

\paragraph{(F2.4) Breakdown of the physical axis $A_{\mathrm{phys}}$.}
$K_2$ requires a new axis with distinct admissible states (e.g.~spatial
dimension, phase angle, field intensity).
If an axis with $|A|<2$ cannot be defined, the continuum fails.



% ================================================================
\subsection{Percolation and Threshold Falsifiability}
% ================================================================

The fundamental transition from $K_1$ to $K_2$ is governed by a
percolation threshold $p_c$ and the dimensional threshold
$\Theta_{\mathrm{dim}}$.

$K_2$ is falsified if:

\paragraph{(F2.5) The critical threshold \texorpdfstring{$p_c$}{p_c} cannot be defined.}
If no mathematically coherent condition separates connected and
disconnected phases, $K_2$ loses its defining structure.

\paragraph{(F2.6) \texorpdfstring{$p_c$}{p_c} is never attainable.}
If the admissible configurations of $K_1$ lie permanently in
$p<p_c$, $K_2$ cannot emerge.

\paragraph{(F2.7) The system overshoots the threshold incoherently.}
If crossing $p_c$ causes divergence or collapse of $\Omega$ instead of
forming a connected phase, the continuum fails.

\paragraph{(F2.8) Dimensional threshold violation.}
If the structural tension of $K_1$ does not reach
\[
T_1 > \Theta_{\mathrm{dim}},
\]
no new axis can form, and $K_2$ cannot be born.

\paragraph{(F2.9) Threshold landscape contradiction.}
If the physical threshold set $\Theta_{\mathrm{phys}}$ cannot be
consistently defined, physical interactions cannot stabilise.



% ================================================================
\subsection{Boundary and Admissibility Falsifiability}
% ================================================================

The physical continuum requires a non-empty admissible region
$\Omega(K_2)$ with a well-defined boundary $\partial\Omega(K_2)$.

$K_2$ is falsified when:

\paragraph{(F2.10) \texorpdfstring{$\Omega(K_2)=\varnothing$}{\Omega(K_2)=\varnothing}.}
If all physically admissible configurations violate thresholds,
admissibility collapses.

\paragraph{(F2.11) \texorpdfstring{$\partial\Omega(K_2)$}{\partial\Omega(K_2)} cannot be defined.}
Boundaries must be set by physical fields, gradients, and potentials.
If these contradict each other, $\partial\Omega$ becomes undefined.

\paragraph{(F2.12) \texorpdfstring{$\partial\Omega(K_2)$}{\partial\Omega(K_2)} contradicts continuity.}
If the boundary cannot be embedded in $\Omega(K_1)$, the transition is
invalid.

\paragraph{(F2.13) Non-physical configurations leak into the admissible
set.}
If configurations violating interaction rules or connectivity are still
admissible, the continuum becomes inconsistent.



% ================================================================
\subsection{Potentials, Flows and Field Falsifiability}
% ================================================================

At $K_2$ physical potentials $P_{\mathrm{phys}}$ and flows
$J_{\mathrm{phys}}$ appear for the first time.

$K_2$ is falsified if:

\paragraph{(F2.14) Physical potentials cannot be defined.}
If no consistent scalar or vector potentials can be assigned, physical
interaction is impossible.

\paragraph{(F2.15) Potentials violate admissibility ranges.}
If $P_{\mathrm{phys}}$ cannot be kept within the allowable region
defined by $\Theta_{\mathrm{phys}}$, the continuum collapses.

\paragraph{(F2.16) Physical flows $J_{\mathrm{phys}}$ cannot be formed.}
Flows must be definable from gradients and potentials.
If no such dynamics exist, $K_2$ fails.

\paragraph{(F2.17) Flow divergence destroys connectivity.}
If flows necessarily break the percolated structure, the continuum
cannot maintain a global state.



% ================================================================
\subsection{Dynamic and Evolutionary Falsifiability}
% ================================================================

The continuum $K_2$ evolves under its physical operator:
\[
K_2(t+dt) = F_2(K_2(t)),
\]
where $F_2$ incorporates physical potentials, flows, and connectivity
constraints.

It is falsified if:

\paragraph{(F2.18) \texorpdfstring{$F_2$}{F_2} cannot be defined.}
If no lawful evolution exists, the continuum is invalid.

\paragraph{(F2.19) \texorpdfstring{$F_2$}{F_2} breaks admissibility.}
If evolution forces the system outside $\Omega(K_2)$ or below $p_c$,
stability is lost.

\paragraph{(F2.20) \texorpdfstring{$F_2$}{F_2} destroys the new axis.}
If evolution eliminates the admissible states of $A_{\mathrm{phys}}$, the
continuum collapses.

\paragraph{(F2.21) No stable attractors exist.}
Physical continuum requires attractors or stable regions.
If all trajectories diverge, $K_2$ is falsified.



% ================================================================
\subsection{Transition and Embedding Falsifiability}
% ================================================================

The transition $K_1 \to K_2$ is governed by the operator $F_{1\to 2}$ and
requires a coherent embedding in the meta-space $M_2$.

$K_2$ is falsified if:

\paragraph{(F2.22) $F_{1\to 2}$ cannot act.}
If $K_1$ cannot generate the new axis or connectivity, the transition is
impossible.

\paragraph{(F2.23) Embedding inconsistency.}
If $K_2$ cannot be embedded in the admissible region of $M_2$, it cannot
exist.

\paragraph{(F2.24) Conflict with physical laws encoded in \texorpdfstring{$M_2$}{M_2}.}
If physical interactions required by $K_2$ contradict the structure of
$M_2$, the continuum fails.



% ================================================================
\subsection{Summary}
% ================================================================

$K_2$ is falsified whenever:

\begin{itemize}
    \item no global connected component forms,
    \item the percolation threshold $p_c$ is unattainable or undefined,
    \item the physical axis $A_{\mathrm{phys}}$ cannot emerge,
    \item $\Omega(K_2)$ or its boundary $\partial\Omega(K_2)$ cannot be
    defined,
    \item physical potentials and flows cannot be introduced,
    \item evolution under $F_2$ is impossible or destroys stability,
    \item the transition $K_1\to K_2$ fails structurally or dynamically.
\end{itemize}

Failure of any of these conditions renders the physical continuum
$K_2$ impossible and blocks the entire physical branch of the OC
hierarchy.
