% ================================================================
% ==== FILE: content/falsifiability/falsifiability_k1.tex
% ================================================================

\section{Falsifiability of $K_1$}
\label{sec:falsifiability-k1}

The continuum $K_1$ is the first non-trivial continuum in the hierarchy
of the Ontology of Continua.  
It introduces a one-dimensional axis of admissible variation,
typically interpreted as the minimal temporal or parametric axis $\tau$,
together with a space of fields on this axis:
\[
K_1 = (X,\tau, A_1, V, \Omega, E, \varphi_0).
\]

Unlike $K_0$, which is purely structural and pre-dynamic, $K_1$ already
possesses:
\begin{itemize}
    \item a topological line $(X,\tau)$,
    \item a field space $V$,
    \item an admissible configuration space $\Omega=C^{0}(X,V)$,
    \item an energy functional $E$ obeying axioms E1–E3,
    \item a non-trivial admissibility region with boundary $\partial\Omega$,
    \item a minimal threshold $\Theta_1$ for allowed configurations,
    \item the trivial cycle $C_{\mathrm{triv}}$ defined over the axis.
\end{itemize}

Falsifiability of $K_1$ concerns the possibility that any of these
structures fail to meet the defining axioms of the continuum.



% ================================================================
\subsection{Structural and Topological Falsifiability}
% ================================================================

The continuum $K_1$ is falsified if its minimal topological and field
structures cannot be coherently defined.  
The key conditions include:

\paragraph{(F1.1) Breakdown of the axis $(X,\tau)$.}
$K_1$ requires a one-dimensional continuum with a valid topology.
If $(X,\tau)$ cannot be defined, is inconsistent, or collapses into a
disconnected or discrete structure incompatible with the axioms, the
continuum fails.

\paragraph{(F1.2) Invalid or ill-defined field space $V$.}
The field space must be a well-defined vector or configuration space;
if $V$ admits contradictions or cannot support continuous fields,
$\Omega$ becomes ill-defined.

\paragraph{(F1.3) Non-existence of the admissible configuration space $\Omega=C^0(X,V)$.}
If continuity cannot be defined on $(X,\tau)$ or if $V$ does not support
continuous maps, then $\Omega=\varnothing$ and $K_1$ is impossible.

\paragraph{(F1.4) Inconsistent admissibility region.}
The admissible region $\Omega_{\mathrm{adm}}$ must satisfy the $K_1$
constraints.
If it cannot be defined or is immediately empty, the continuum fails.



% ================================================================
\subsection{Energy Functional and Threshold Falsifiability}
% ================================================================

The energy functional
\[
E : \Omega \to \mathbb{R}
\]
is essential for defining admissibility at $K_1$, together with its axioms
(E1: boundedness below, E2: locality, E3: stability).

$K_1$ is falsified when:

\paragraph{(F1.5) $E$ violates boundedness below (E1).}
If $E$ has no lower bound, the continuum becomes dynamically unstable.

\paragraph{(F1.6) $E$ violates locality (E2).}
Non-local energy at $K_1$ contradicts the definition of the continuum and
invalidates its structure.

\paragraph{(F1.7) $E$ violates stability (E3).}
If small perturbations lead to divergent energy without a stabilising
mechanism, $K_1$ collapses.

\paragraph{(F1.8) Violation of the threshold $\Theta_1$.}
If no configuration satisfies
\[
E(\varphi) \le \Theta_1,
\]
or if $\Theta_1$ cannot be defined consistently, $K_1$ is impossible.



% ================================================================
\subsection{Boundary and Region Falsifiability}
% ================================================================

The admissible region $\Omega$ has a boundary $\partial\Omega$ defined by
energy and continuity constraints.
$K_1$ is falsified if:

\begin{itemize}
    \item $\partial\Omega$ cannot be defined,
    \item $\partial\Omega$ contradicts the topology $(X,\tau)$,
    \item $\Omega$ collapses to the empty set,
    \item admissible configurations violate the $K_1$ axioms.
\end{itemize}

The boundary is essential for the continuum's existence; without a valid
boundary, the continuum has no admissible states.



% ================================================================
\subsection{Cycle and Evolution Falsifiability}
% ================================================================

Even though $K_1$ is minimally dynamic, it must support the trivial cycle
$C_{\mathrm{triv}}$ and the minimal action of the operator $F_1$,
defined by
\[
K_1(t+dt) = F_1(K_1(t)).
\]

The continuum is falsified if:

\paragraph{(F1.9) The trivial cycle cannot be defined.}
If $C_{\mathrm{triv}}$ fails to exist, continuity along the axis is
impossible.

\paragraph{(F1.10) The operator $F_1$ cannot act consistently.}
If evolution under $F_1$ breaks admissibility or contradicts the axioms,
$K_1$ collapses.

\paragraph{(F1.11) No stable configuration $\varphi_0$.}
If there exists no baseline configuration $\varphi_0$ satisfying
continuity and threshold conditions, the continuum is inconsistent.



% ================================================================
\subsection{Embedding-Space and Transition Falsifiability}
% ================================================================

The transition from $K_0$ to $K_1$ is governed by the operator
$\Psi_{0\to 1}$ and requires the existence of the one-dimensional axis as
a new admissible dimension.

$K_1$ is falsified when:

\paragraph{(F1.12) $\Psi_{0\to 1}$ cannot act.}
If the embedding space of $K_0$ cannot support a one-dimensional axis,
no $K_1$ can arise.

\paragraph{(F1.13) Dimensional threshold failure.}
If the structural tension at $K_0$ does not reach
\[
T_0 > \Theta_{\mathrm{dim}},
\]
the new axis cannot emerge.

\paragraph{(F1.14) Embedding contradiction.}
If $K_1$ cannot be embedded within the admissible meta-space specified by
$M_1$, the continuum is impossible.



% ================================================================
\subsection{Summary}
% ================================================================

$K_1$ is falsified whenever:

\begin{itemize}
    \item its minimal topology $(X,\tau)$ is inconsistent,
    \item its field space $V$ is invalid or contradictory,
    \item the configuration space $\Omega$ cannot be defined,
    \item the energy functional $E$ violates axioms E1–E3,
    \item the threshold $\Theta_1$ cannot be satisfied,
    \item the boundary $\partial\Omega$ breaks consistency,
    \item the trivial cycle $C_{\mathrm{triv}}$ cannot exist,
    \item the evolution operator $F_1$ cannot act coherently,
    \item the transition from $K_0$ to $K_1$ fails structurally.
\end{itemize}

Any such violation renders the continuum $K_1$ impossible and blocks the
entire emergence of higher continua $K_2$–$K_{12}$.
