% ================================================================
% ==== FILE: content/falsifiability/falsifiability_k8.tex
% ================================================================

\section{Falsifiability of \texorpdfstring{$K_8$}{K_8}}
\label{sec:falsifiability-k8}

The continuum $K_8$ describes civilisation-level organisation:
symbolic systems, cumulative knowledge, institutional memory,
technological infrastructure, written communication,
and long-range coordination across large populations and long timescales.

A valid $K_8$ requires:

\begin{itemize}
    \item axes $A_8$: writing, symbolic systems, knowledge structures,
          large-scale institutions, technology, cultural reproduction;
    \item potentials $P_8$: epistemic potential, symbolic-energy potential,
          technological potential, institutional stability,
          cultural coherence, long-range coordination potential;
    \item thresholds $\Theta_8$: epistemic threshold,
          cultural-coherence threshold,
          institutional-stability threshold,
          technological-sufficiency threshold,
          complexity-control threshold;
    \item flows $J_8$: knowledge flows, symbolic flows,
          technological flows, institutional flows,
          long-range communication flows;
    \item cycles $C_8$: knowledge-cycle, technological-cycle,
          institutional-cycle, cultural-cycle,
          innovation-cycle, archival-cycle;
    \item stable civilisation space $\Omega(K_8)$ and boundary $\partial\Omega(K_8)$;
    \item bounded structural tension $T_8$;
    \item non-zero continuumness $k_8>0$.
\end{itemize}

Violation of any of these leads to falsification of $K_8$.



% ================================================================
\subsection{Falsifiability via Symbolic Systems and Writing}
% ================================================================

Symbolic systems and writing are defining axes of $K_8$.
They enable cumulative culture, long-range memory,
and expansion of social organisation beyond individual cognition.

$K_8$ is falsified when:

\paragraph{(F8.1) No stable symbolic system.}
If symbols cannot stabilise or be interpreted consistently,
cumulative knowledge cannot form.

\paragraph{(F8.2) Writing instability.}
If written records degrade faster than they are reproduced,
the civilisation cannot preserve information.

\paragraph{(F8.3) Loss of symbolic coherence.}
If symbolic systems fragment beyond $\Theta_{\mathrm{symbol}}$,
$\Omega_8$ loses connectedness.

\paragraph{(F8.4) Symbolic overload.}
If symbolic complexity grows beyond $\Theta_{\mathrm{symbolic\_load}}$,
interpretability collapses.

\paragraph{(F8.5) Divergent symbolic tension.}
If symbolic demands exceed capacity of institutions and cognition,
$T_8$ diverges, destroying civilisation-level structure.



% ================================================================
\subsection{Falsifiability via Knowledge Systems and Epistemic Structure}
% ================================================================

Knowledge systems define civilisation-level memory.

$K_8$ is falsified when:

\paragraph{(F8.6) Epistemic threshold not met.}
If epistemic potentials are insufficient
to encode, preserve, and reproduce complex knowledge,
civilisation cannot emerge.

\paragraph{(F8.7) Epistemic collapse.}
If knowledge decays faster than it accumulates,
the system reverts to $K_7$.

\paragraph{(F8.8) Epistemic fragmentation.}
If incompatible knowledge clusters exceed
$\Theta_{\mathrm{epistemic-frag}}$,
global coherence collapses.

\paragraph{(F8.9) Breakdown of scientific/institutional memory.}
Failure of science, archives, schools, universities,
or epistemic institutions destroys $\Omega_8$.

\paragraph{(F8.10) Cognitive–epistemic mismatch.}
If cognitive capacities of individuals ($K_6$)
cannot support epistemic demands of $K_8$,
civilisation becomes unstable.



% ================================================================
\subsection{Falsifiability via Institutions and Governance}
% ================================================================

Civilisations require stable large-scale institutions.

$K_8$ is falsified when:

\paragraph{(F8.11) Institutional instability.}
If institutions oscillate beyond $\Theta_{\mathrm{instab}}$,
they cannot maintain civilisation-scale coherence.

\paragraph{(F8.12) Collapse of coordination institutions.}
If coordination, enforcement, or governance institutions fail,
long-range organisation breaks down.

\paragraph{(F8.13) Institutional–cultural mismatch.}
If institutions contradict cultural norms beyond tolerance,
$T_8$ diverges.

\paragraph{(F8.14) Bureaucratic overload.}
If administrative complexity exceeds $\Theta_{\mathrm{bureaucratic}}$,
institutional flows $J_8$ stall.

\paragraph{(F8.15) Institutional memory decay.}
If institutions lose continuity, civilisation loses stability.



% ================================================================
\subsection{Falsifiability via Technology and Infrastructure}
% ================================================================

Technological systems in $K_8$ create new degrees of freedom
for resource flow, communication, and symbolic reproduction.

$K_8$ is falsified when:

\paragraph{(F8.16) Technological sufficiency threshold not met.}
If technologies required for knowledge reproduction do not exist,
civilisation cannot persist.

\paragraph{(F8.17) Infrastructure collapse.}
If communication, energy, or transport infrastructure collapses,
global cycles $C_8$ break.

\paragraph{(F8.18) Tech–institution mismatch.}
If institutions cannot regulate or integrate technological change,
stability fails.

\paragraph{(F8.19) Negative technological externalities.}
If technological side-effects exceed $\Theta_{\mathrm{tech\_damage}}$,
structure becomes unstable.

\paragraph{(F8.20) Overacceleration.}
If technological development grows faster than the adaptation capacity
of institutions and culture, civilisation collapses.



% ================================================================
\subsection{Falsifiability via Cultural Dynamics}
% ================================================================

Culture ensures long-term reproduction of civilisation patterns.

$K_8$ is falsified when:

\paragraph{(F8.21) Cultural incoherence.}
If cultural potentials diverge beyond $\Theta_{\mathrm{culture}}$,
civilisation loses identity and stability.

\paragraph{(F8.22) Cultural stagnation.}
If cultural reproduction fails (loss of traditions, language, norms),
institutional stability collapses.

\paragraph{(F8.23) Cultural fragmentation.}
If groups split into incompatible cultural clusters,
$\Omega_8$ becomes disconnected.

\paragraph{(F8.24) Cultural–symbolic divergence.}
If cultural norms contradict symbolic systems (writing, science, law),
the system becomes unstable.

\paragraph{(F8.25) Cultural entropy.}
Excessive randomness in cultural reproduction
destroys large-scale coherence.



% ================================================================
\subsection{Falsifiability via Civilisational Flows \texorpdfstring{$J_8$}{J_8}}
% ================================================================

Civilisation requires stable flows of information, resources,
knowledge, symbols, and institutional authority.

$K_8$ is falsified when:

\paragraph{(F8.26) Breakdown of long-range communication.}
If communication networks fail,
civilisation-scale organisation collapses.

\paragraph{(F8.27) Resource flow collapse.}
If energy, food, or material flows collapse,
institutions break down.

\paragraph{(F8.28) Knowledge-flow bottlenecks.}
If knowledge cannot propagate,
the innovation cycle fails.

\paragraph{(F8.29) Symbolic-flow imbalance.}
If symbolic flows overload cultural or institutional systems,
$T_8$ increases beyond stability.

\paragraph{(F8.30) Institutional-flow disruption.}
If authority or governance flows collapse,
$\Omega_8$ disintegrates.



% ================================================================
\subsection{Falsifiability via Civilisational Cycles \texorpdfstring{$C_8$}{C_8}}
% ================================================================

$K_8$ depends on cycles:

\[
C_8 = \{
C_{\mathrm{knowledge}},
C_{\mathrm{technology}},
C_{\mathrm{institution}},
C_{\mathrm{culture}},
C_{\mathrm{innovation}},
C_{\mathrm{archive}}
\}.
\]

$K_8$ is falsified when:

\paragraph{(F8.31) Knowledge-cycle collapse.}
If knowledge cannot be preserved, tested, and transmitted,  
civilisation regresses to $K_7$.

\paragraph{(F8.32) Institutional-cycle failure.}
If institutions cannot reproduce themselves,
long-term coherence is lost.

\paragraph{(F8.33) Innovation-cycle breakdown.}
If innovation cannot be sustained,
technological stagnation collapses $P_8$.

\paragraph{(F8.34) Archival-cycle failure.}
If archives cannot preserve symbolic systems,
$\Omega_8$ collapses.

\paragraph{(F8.35) Culture-cycle failure.}
If culture cannot reproduce itself,
civilisation dissolves.



% ================================================================
\subsection{Falsifiability via Structural Tension \texorpdfstring{$T_8$}{T_8}}
% ================================================================

Structural tension at $K_8$ includes:

- epistemic tension,
- symbolic tension,
- institutional tension,
- cultural tension,
- technological tension.

$K_8$ is falsified when:

\paragraph{(F8.36) \texorpdfstring{$T_8$}{T_8} exceeds collapse threshold.}
Civilisation cannot stabilise under excessive internal stress.

\paragraph{(F8.37) Multi-axis tension divergence.}
If epistemic, institutional, and technological tensions diverge simultaneously,
collapse is unavoidable.

\paragraph{(F8.38) Boundary instability.}
If $\partial\Omega_8$ (geographical, cultural, or epistemic boundaries)
destabilises, the continuum dies.

\paragraph{(F8.39) Complexity overload.}
If systemic complexity exceeds $\Theta_{\mathrm{complexity}}$,
institutions and culture cannot control it.

\paragraph{(F8.40) System-wide resonance.}
Synchronised failures across $A_8$ axes destroy $k_8$.



% ================================================================
\subsection{Falsifiability of Transition \texorpdfstring{$K_7 \to K_8$}{K_7 \to K_8}}
% ================================================================

The transition $\Psi_{7\to 8}$ requires:

\begin{itemize}
    \item emergence of stable writing and symbolic systems,
    \item scaling of institutions from group-level to civilisation-level,
    \item long-range communication and resource flows,
    \item cumulative culture and knowledge storage,
    \item technological and epistemic support.
\end{itemize}

The transition fails when:

\paragraph{(F8.41) No writing emergence.}
Without writing or symbolic storage,
civilisation cannot form.

\paragraph{(F8.42) Failed scaling of institutions.}
If institutions cannot scale beyond group-level,
$K_8$ cannot stabilise.

\paragraph{(F8.43) Insufficient technological base.}
If technologies enabling symbolic reproduction do not exist,
the transition stalls.

\paragraph{(F8.44) Cultural incoherence at birth.}
If culture cannot support symbolic systems and institutions,
$\Omega_8$ cannot form.

\paragraph{(F8.45) Collapse of long-range flows.}
If long-distance flows $J_8$ cannot stabilise,
the system remains at $K_7$.



% ================================================================
\subsection{Summary}
% ================================================================

$K_8$ is falsified if civilisation-level symbolic, epistemic,
institutional, technological, cultural, or flow structures
cannot stabilise or maintain bounded tension.

If any of these conditions fail,
the civilisation continuum collapses and $K_8$ ceases to exist.
