% ================================================================
% ==== FILE: content/falsifiability/falsifiability_k0.tex
% ================================================================

\section{Falsifiability of $K_0$}
\label{sec:falsifiability-k0}

The continuum $K_0$ represents the minimal structural substrate of the
Ontology of Continua.
It is defined as a pre-dynamic object
\[
K_0 = (S,\Delta,\mathcal{C}),
\]
with no time, no flows, no potentials, no thresholds beyond the minimal
existence bound $\Theta_0=\varepsilon>0$, and a trivial cycle
$C_{\mathrm{triv}}$.
Falsifiability of $K_0$ therefore tests the logical and structural
consistency of the most primitive form of continuumhood.

Despite its simplicity, $K_0$ imposes strict constraints.
Any violation of its axioms is sufficient to falsify not only $K_0$
but also all higher continua, because the full hierarchy
$K_1$–$K_{12}$ is built upon the existence of a well-defined $K_0$.



% ================================================================
\subsection{Fundamental Falsifiability Principles for $K_0$}
% ================================================================

The falsification of $K_0$ occurs when the structural triple 
$(S,\Delta,\mathcal{C})$ fails to satisfy the defining axioms:

\begin{itemize}
    \item \textbf{Axioms D1–D4:}  
          governing the set $S$ of primitive distinctions and the
          minimal structural properties of $\Delta$.

    \item \textbf{Axioms C1–C4$^\prime$:}  
          governing the construction and compatibility of the
          composition operator $\mathcal{C}$.

    \item \textbf{Axiom 0.1 (Existence):}  
          which provides the minimal condition for a continuum to exist.

    \item \textbf{Minimal threshold:}  
          $\Theta_0 = \varepsilon > 0$, ensuring non-degeneracy.
\end{itemize}

Thus, $K_0$ is falsified whenever the primitive combinatorial structure
fails to meet these requirements.



% ================================================================
\subsection{Structural Falsifiability Conditions}
% ================================================================

Several structural incompatibilities can invalidate $K_0$:

\paragraph{(F0.1) Undefined or incoherent $S$.}  
If the primitive set $S$ of distinctions is empty or logically 
inconsistent, no continuum can be constructed.

\paragraph{(F0.2) Breakdown of $\Delta$.}  
If the difference relation $\Delta$ cannot be defined such that it 
respects axioms D1–D4 (irreflexivity, minimal consistency, closure
under composition), then $K_0$ is impossible.

\paragraph{(F0.3) Inconsistent composition $\mathcal{C}$.}  
If $\mathcal{C}$ cannot produce valid composite distinctions or violates 
C1–C4$^\prime$, structural coherence is lost.

\paragraph{(F0.4) Violation of the minimal existence threshold.}  
If under the primitive structure the threshold
\[
\Theta_0 = \varepsilon>0
\]
cannot be maintained (i.e.\ the structure collapses to the empty set),
$K_0$ cannot exist.

\paragraph{(F0.5) Lack of a trivial cycle.}  
Even at $K_0$, the trivial cycle $C_{\mathrm{triv}}$ must exist.
A failure to define this cycle violates the minimal continuity condition.

\paragraph{(F0.6) Contradiction in structural pairs.}  
If the structural pair $(S,\Delta)$ cannot exist without contradiction
(violating the ``structural pair axiom''), $K_0$ is falsified.



% ================================================================
\subsection{Boundary and Region Falsifiability}
% ================================================================

Although $K_0$ has no dynamical boundary in the usual sense,
it still requires a well-defined primitive boundary of admissible 
configurations:
\[
\partial\Omega(K_0) = \{ \text{distinctions permitted by } S,\Delta,\mathcal{C} \}.
\]

$K_0$ is falsified if:

\begin{itemize}
    \item $\partial\Omega$ cannot be defined (the primitive region has no
          consistent boundary),
    \item $\Omega(K_0)=\varnothing$,
    \item the boundary admits contradictions in distinction composition.
\end{itemize}

In other words, if the primitive “space of distinctions” cannot be
constructed, no continuum can exist at all.



% ================================================================
\subsection{Collapse and Non-Existence at $K_0$}
% ================================================================

Since $K_0$ has no dynamics, collapse is strictly structural rather than
temporal.
The continuum $K_0$ fails when it cannot satisfy its definition in a
single static configuration.

This occurs when:

\begin{enumerate}
    \item $S$ degenerates into a contradictory or empty set;
    \item $\Delta$ cannot be defined coherently;
    \item $\mathcal{C}$ produces contradictions or cannot operate;
    \item $\Theta_0$ cannot be respected;
    \item $C_{\mathrm{triv}}$ cannot be defined.
\end{enumerate}

Any such violation implies the total non-existence of $K_0$ and therefore 
renders all higher continua impossible.



% ================================================================
\subsection{Embedding-Space Constraints for $K_0$}
% ================================================================

Although $K_0$ has no parent continuum, it is still embedded in the 
minimal admissible meta-space that supports primitive distinction
structures.
Falsifiability via embedding arises when:

\begin{itemize}
    \item the meta-space cannot host any non-degenerate distinction set,
    \item the assumed primitive distinctions require an impossible 
          background structure,
    \item $\Omega(K_0)$ cannot be embedded into this minimal meta-space.
\end{itemize}

These conditions guarantee that $K_0$ is not merely a logical
construction but must be realisable within an admissible structural 
environment.



% ================================================================
\subsection{Summary}
% ================================================================

$K_0$ is falsified purely by structural contradictions.
Because $K_0$ contains no time, no potentials, no flows, and no
dynamics, every falsification is immediate and global.
A single violation of:

\begin{itemize}
    \item the primitive distinction set $S$,
    \item the difference relation $\Delta$,
    \item the composition operation $\mathcal{C}$,
    \item the minimal threshold $\Theta_0$,
    \item the existence of $C_{\mathrm{triv}}$,
    \item or the coherence of $\partial\Omega(K_0)$
\end{itemize}

invalidates $K_0$ entirely.

Because every higher continuum $K_x$ requires a valid $K_0$,
falsifying $K_0$ collapses the entire hierarchy.
