% ================================================================
% ==== FILE: content/k_levels/k0.tex
% ================================================================

% ==============================
%  Ontology of Continua — Core
%  K-level Module: K0
%  File: content/k_levels/k0.tex
%  Status: FULLY DEFINED
% ==============================

\subsubsection{\texorpdfstring{$K_0$}{K_0} Overview}
\label{sec:k0-overview}

Level $K_0$ represents the pre-structural foundation of all continua.
It is not a physical or geometric domain but the logical substrate that
defines the \emph{conditions of possibility} for any continuum $K_x$.
It provides the metapressure, admissibility constraints, logical
superiority (Axiom~0.3), and the universal rules that govern which
structures may exist.

$K_0$ does not itself possess the conventional axes or geometric
dimensions.  
Instead, it specifies:
\begin{itemize}
    \item the meta-laws $\mathcal{L}$ determining structural coherence,
    \item the possibility space $\mathcal{P}$ for all continua,
    \item the metapressure driving emergence of differences,
    \item the universal thresholds that define admissibility and collapse,
    \item the recursion rules $K_x \mapsto M_x \mapsto K_{x+1}$.
\end{itemize}

Its primary function is \textbf{logical governance} of all deeper
levels.

% ================================================================
\subsubsection{State Space \texorpdfstring{$\Omega(K_0)$}{\Omega(K_0)}}

$\Omega(K_0)$ consists of all admissible configurations of the pair
\[
(\mathcal{P}, \mathcal{L}),
\]
where $\mathcal{P}$ is the meta-domain of possible structures and
$\mathcal{L}$ is the system of meta-laws that enforce coherence.

Because $K_0$ is pre-geometric, $\Omega(K_0)$ is not a metric space but
an abstract set with the following properties:
\begin{enumerate}
    \item $\Omega(K_0)$ contains all possible specifications of
          structural constraints that allow continua to exist.
    \item $\Omega(K_0)$ has no internal topology unless induced by
          constraints in $\mathcal{L}$.
    \item Every $\Omega(K_x)$ for $x \ge 1$ is a \emph{subset} of
          $\Omega(K_0)$:
          \[
              \Omega(K_x) \subseteq \Omega(K_0).
          \]
\end{enumerate}

There is no continuous evolution inside $\Omega(K_0)$; it defines
\emph{conditions}, not dynamical trajectories.

% ================================================================
\subsubsection{Boundary \texorpdfstring{$\partial\Omega(K_0)$}{\partial\Omega(K_0)}}

The boundary of $K_0$ consists of all meta-configurations that violate
internal coherence conditions of the meta-laws $\mathcal{L}$.

A configuration $(\mathcal{P},\mathcal{L})$ lies on $\partial\Omega(K_0)$ if:
\begin{itemize}
    \item the laws contradict each other,
    \item admissibility collapses ($\mathcal{P}$ becomes empty),
    \item recursion rules become undefined,
    \item structural consistency conditions cannot be maintained.
\end{itemize}

Crossing $\partial\Omega(K_0)$ corresponds to the logical impossibility 
of any continuum.

% ================================================================
\subsubsection{Axes \texorpdfstring{$A(K_0)$}{A(K_0)}}

$K_0$ has no geometric axes. Instead it contains:
\begin{itemize}
    \item abstract axes of \emph{possibility},
    \item admissibility axes governing which structures can be produced,
    \item meta-axes that define the logical dependencies among levels.
\end{itemize}

In formal terms:
\[
A(K_0) = \{ \text{dimensions of admissible structural variation} \}.
\]

These are the ultimate sources of all later axes:
\[
A(K_x) \subseteq A(K_0).
\]

% ================================================================
\subsubsection{Potentials \texorpdfstring{$P(K_0)$}{P(K_0)}}

Potentials at $K_0$ describe meta-conditions that push continua toward
structure:
\[
P(K_0) = \{ \text{pressures, constraints, logical gradients} \}.
\]

They include:
\begin{itemize}
    \item metapressure (Axiom 0.z),
    \item pressure toward structuralisation,
    \item pressure toward compatible differences,
    \item pressure toward nonempty admissible spaces.
\end{itemize}

These potentials are not numeric energies but logical intensities.

% ================================================================
\subsubsection{Thresholds \texorpdfstring{$\Theta(K_0)$}{\Theta(K_0)}}

Thresholds at $K_0$ correspond to logical break-points:
\[
\Theta(K_0) = \{ \Theta_{\mathrm{coh}}, \Theta_{\mathrm{adm}}, 
                 \Theta_{\mathrm{meta}} \}.
\]

Examples:
\begin{itemize}
    \item $\Theta_{\mathrm{coh}}$: minimal coherence of meta-laws,
    \item $\Theta_{\mathrm{adm}}$: minimal non-empty possibility space,
    \item $\Theta_{\mathrm{meta}}$: minimal support for recursion.
\end{itemize}

If any threshold is violated, no continuum can exist at any $K_x$.

% ================================================================
\subsubsection{Flows \texorpdfstring{$J(K_0)$}{J(K_0)}}

There are no physical flows at $K_0$.  
$J(K_0)$ instead contains:
\begin{itemize}
    \item flows of admissibility,
    \item flows of logical dependence,
    \item propagation of constraints,
    \item meta-flows describing which structures can generate others.
\end{itemize}

These flows shape the space of possible continua but do not correspond
to time evolution.

% ================================================================
\subsubsection{Cycles \texorpdfstring{$C(K_0)$}{C(K_0)}}

Cycles at $K_0$ correspond to:
\begin{itemize}
    \item fixed-point consistency cycles,
    \item meta-stability loops of admissibility,
    \item self-referential closure cycles in $\mathcal{L}$.
\end{itemize}

A cycle exists if:
\[
\Pi(C) > \Theta_{\mathrm{coh}},
\]
where $\Pi(C)$ is the ``power'' of the meta-cycle (stability of recursive
closure).

There are no temporal cycles; these are purely logical cycles.

% ================================================================
\subsubsection{Time \texorpdfstring{$\tau(K_0)$}{\tau(K_0)}}

Time does not exist at $K_0$.

Theorem “Emergence of Time from C-cycles” (Physics Run):

Time emerges when $\Omega(K_2)$ contains a non-degradable cycle
$C$ with:
\[
\Pi(C) > \Theta_{\mathrm{time}}.
\]

Thus:
\[
\tau(K_0) \;\text{undefined}.
\]

% ================================================================
\subsubsection{Continuumness \texorpdfstring{$k(K_0)$}{k(K_0)}}

Continuumness measures the admissibility of all continua:
\[
k(K_0) = 
H(\Omega(K_0)) \cdot
k_A(K_0) \cdot
k_{\mathrm{coh}}(K_0),
\]
where:
\begin{itemize}
    \item $H(\Omega(K_0)) = 1$ if $\Omega(K_0)\neq\emptyset$, else $0$,
    \item $k_A(K_0)$ measures the richness of admissible axes,
    \item $k_{\mathrm{coh}}$ measures meta-law coherence.
\end{itemize}

If $k(K_0)=0$, no continuum at any higher level can exist.

% ================================================================
\subsubsection{Structural Tension \texorpdfstring{$T(K_0)$}{T(K_0)}}

Structural tension is:
\[
T(K_0) = T(\mathcal{P}, \mathcal{L}),
\]
representing the degree of incompatibility within meta-laws or between
meta-laws and admissible structures.

High tension may:
\begin{itemize}
    \item reduce admissible sets,
    \item produce incompatible branches,
    \item collapse $\Omega(K_0)$.
\end{itemize}

% ================================================================
\subsubsection{Energy \texorpdfstring{$E(K_0)$}{E(K_0)}}

There is no physical energy at $K_0$.  
Instead:
\[
E(K_0) = E_{\mathrm{meta}},
\]
a formal measure of internal support for consistent structures:
\[
E_{\mathrm{meta}} \propto \text{coherence of }\mathcal{L}.
\]

% ================================================================
\subsubsection{Operators on \texorpdfstring{$K_0$}{K_0} (\texorpdfstring{$\Psi$}{\Psi}, \texorpdfstring{$\Phi$}{\Phi}, \texorpdfstring{$\Lambda$}{\Lambda}, \texorpdfstring{$U$}{U}, \texorpdfstring{$\Chi$}{\Chi})}

Operators at $K_0$ govern admissibility and recursion:

\begin{itemize}
    \item $\Psi$: operator of level creation  
          \[
          \Psi: K_x \mapsto M_{x+1}
          \]
    \item $\Phi$: operator of meta-space evolution  
          \[
          \Phi: M_x \mapsto M_x(t+dt)
          \]
    \item $\Lambda$: constraint operator enforcing $\mathcal{L}$
    \item $U$: operator of structural unification
    \item $\Chi$: operator of logical branching (Axiom~21)
\end{itemize}

These operators define the backbone of vertical recursion.

% ================================================================
\subsubsection{Processes on \texorpdfstring{$K_0$}{K_0}}

Key processes:
\begin{itemize}
    \item logical unification,
    \item admissibility expansion/contraction,
    \item meta-stabilisation of recursive definitions,
    \item branching of possibility spaces,
    \item collapse of incompatible meta-structures.
\end{itemize}

% ================================================================
\subsubsection{Predictions for \texorpdfstring{$K_0$}{K_0}}

The theory predicts:
\begin{itemize}
    \item no observable dynamics at $K_0$,
    \item universality of thresholds for all levels,
    \item inevitability of dimensional growth under metapressure,
    \item existence of meta-ontological branching structures.
\end{itemize}

% ================================================================
\subsubsection{Experiments for \texorpdfstring{$K_0$}{K_0}}

$K_0$ cannot be experimentally probed directly.  
Indirect tests:
\begin{itemize}
    \item universality of threshold behaviour across physical,
          chemical, biological, cognitive and social systems,
    \item ubiquity of structural recursion patterns,
    \item falsifiability tests of emergent time (Physics Run),
    \item stability of branching structures (K$_{11}$–K$_{12}$).
\end{itemize}

% ================================================================
\subsubsection{Collapse and Death of \texorpdfstring{$K_0$}{K_0}}

Collapse occurs if:
\[
\Omega(K_0) = \varnothing.
\]

This corresponds to:
\begin{itemize}
    \item contradictory meta-laws,
    \item logically inconsistent possibility space,
    \item unsupportable recursion $\Psi$ or $\Phi$,
    \item violation of $\Theta_{\mathrm{coh}}$ or $\Theta_{\mathrm{adm}}$.
\end{itemize}

If $K_0$ collapses, all continua collapse simultaneously.

% ================================================================
\subsubsection{Falsifiability of \texorpdfstring{$K_0$}{K_0}}

The following predictions are falsifiable:
\begin{enumerate}
    \item Universal presence of thresholds $\Theta$ across all levels.
    \item Irreversibility of dimensional growth (Theorem~1).
    \item Emergence of time only at $K_2$ through cycles.
    \item Existence of universal recursion $K_x\to M_x\to K_{x+1}$.
\end{enumerate}

Failure of any of these behaviours empirically falsifies $K_0$.

% ================================================================
\subsubsection{Branching / Ontological Position of \texorpdfstring{$K_0$}{K_0}}

Based on Axiom~21 (Ontological Branching), $K_0$ is the root node of the
ontological branching tree:
\[
K_0 \rightarrow K_1 \rightarrow K_2 \rightarrow \dots \rightarrow K_{12}.
\]

All branches must remain compatible with $\mathcal{L}$; incompatible
branches die through collapse of admissibility.

% ================================================================
\subsubsection{Relation to M-spaces}

$K_0$ defines the admissibility and structure of all meta-spaces:
\[
M_x \subseteq \Omega(K_0).
\]

Conversely:
\[
K_x \subseteq M_x,
\qquad
M_x \subseteq \Omega(K_{x+1}),
\]
which makes $K_0$ the single unifying superspace for the entire
hierarchy.

