% ================================================================
% ==== FILE: content/k_levels/k10.tex
% ================================================================

% ==============================
%  Ontology of Continua — Core
%  K-level Module: K10 (Metatheoretical Continua)
%  FULL MODULE — FINAL
% ==============================

\subsubsection{$K_{10}$ Overview}
\label{sec:k10-overview}

$K_{10}$ is the level of \emph{metatheoretical continua}: systems capable of
representing, evaluating and transforming whole classes of theories ($K_9$)
including themselves. These continua encode:
\begin{itemize}
    \item metatheoretical axioms,
    \item meta-rules for theory comparison and evaluation,
    \item meta-models describing structures of theories,
    \item universal operators governing theoretical evolution,
    \item recursive self-reference and structural fixed points,
    \item global coherence constraints across multiple theories.
\end{itemize}

Distinctive features:
\begin{itemize}
    \item full self-referential capacity,
    \item global integrative power across lower levels,
    \item meta-consistency constraints,
    \item ability to detect contradictions in $K_9$ theories,
    \item universalization of thresholds and operators,
    \item preparation of the space for $K_{11}$ (structural recursion) via $\Psi_{10\to11}$.
\end{itemize}

% ================================================================
\subsubsection{State Space $\Omega(K_{10})$}

\[
\Omega(K_{10})=
\{
A_{\mathrm{meta}},\;
R_{\mathrm{meta}},\;
M_{\mathrm{meta}},\;
C_{\mathrm{cross}},\;
E_{\mathrm{eval}},\;
S_{\mathrm{self}},\;
\mu^{(10)}_t
\}.
\]

Components:
\begin{itemize}
    \item $A_{\mathrm{meta}}$ — meta-axioms (statements about theories),
    \item $R_{\mathrm{meta}}$ — meta-rules for theory modification,
    \item $M_{\mathrm{meta}}$ — models of theoretical spaces,
    \item $C_{\mathrm{cross}}$ — cross-theory comparison structures,
    \item $E_{\mathrm{eval}}$ — evaluation operators,
    \item $S_{\mathrm{self}}$ — self-representational state,
    \item $\mu^{(10)}_t$ — distribution over metatheoretical configurations.
\end{itemize}

% ================================================================
\subsubsection{Boundary $\partial\Omega(K_{10})$}

Boundary conditions occur when:
\begin{itemize}
    \item self-reference becomes paradoxical,
    \item meta-axioms contradict each other,
    \item evaluation operators lose definability,
    \item meta-rules diverge (no fixed point),
    \item global coherence across $K_9$ theories collapses,
    \item metatheory becomes incompatible with $M_{10}$ constraints,
    \item recursion ($K_{10}\to K_{10}$) becomes ill-formed.
\end{itemize}

Crossing $\partial\Omega(K_{10})$ destroys metatheory as a continuum.

% ================================================================
\subsubsection{Axes $A(K_{10})$}

\[
A(K_{10})=
\{
A_{\mathrm{metaaxiom}},\;
A_{\mathrm{metarule}},\;
A_{\mathrm{metamodel}},\;
A_{\mathrm{evaluation}},\;
A_{\mathrm{coherence}},\;
A_{\mathrm{self}}
\}.
\]

Interpretation:
\begin{itemize}
    \item $A_{\mathrm{metaaxiom}}$: variation of meta-level premises,
    \item $A_{\mathrm{metarule}}$: rules that transform whole theories,
    \item $A_{\mathrm{metamodel}}$: structures modeling $K_9$ landscapes,
    \item $A_{\mathrm{evaluation}}$: global assessment axes,
    \item $A_{\mathrm{coherence}}$: coherence among theories,
    \item $A_{\mathrm{self}}$: depth of self-representation and recursion.
\end{itemize}

% ================================================================
\subsubsection{Potentials $P(K_{10})$}

\[
P(K_{10})=
\{
P_{\mathrm{metaconsistency}},\;
P_{\mathrm{global}},\;
P_{\mathrm{unification}},\;
P_{\mathrm{recursion}},\;
P_{\mathrm{reflection}},\;
P_{\mathrm{selffix}}
\}.
\]

Potentials encode:
\begin{itemize}
    \item meta-consistency (absence of contradictions across theories),
    \item global integrative potential (unifying frameworks),
    \item unification power of meta-structures,
    \item recursive stability,
    \item reflective depth,
    \item existence of self-fixing points ($K_{10}$ describing itself without contradiction).
\end{itemize}

% ================================================================
\subsubsection{Thresholds $\Theta(K_{10})$}

\[
\Theta(K_{10})=
\{
\Theta_{\mathrm{metaconsistency}},\;
\Theta_{\mathrm{coherence}},\;
\Theta_{\mathrm{selfref}},\;
\Theta_{\mathrm{recursion}},\;
\Theta_{\mathrm{unification}},\;
\Theta_{\mathrm{collapse}}
\}.
\]

Descriptions:
\begin{itemize}
    \item $\Theta_{\mathrm{metaconsistency}}$: minimal consistency across theories,
    \item $\Theta_{\mathrm{coherence}}$: global coherence threshold,
    \item $\Theta_{\mathrm{selfref}}$: threshold for safe self-reference,
    \item $\Theta_{\mathrm{recursion}}$: stability under recursive application,
    \item $\Theta_{\mathrm{unification}}$: threshold for unifying divergent theories,
    \item $\Theta_{\mathrm{collapse}}$: metatheoretical failure point.
\end{itemize}

% ================================================================
\subsubsection{Flows $J(K_{10})$}

\begin{itemize}
    \item $J_{\mathrm{metaaxiom}}$: evolution of meta-axioms,
    \item $J_{\mathrm{metarule}}$: transformation of meta-rules,
    \item $J_{\mathrm{metamodel}}$: dynamics of meta-models,
    \item $J_{\mathrm{cross}}$: flow of cross-theory comparisons,
    \item $J_{\mathrm{eval}}$: evolution of evaluation operators,
    \item $J_{\mathrm{self}}$: dynamics of self-representation,
    \item $J_{9\to10}$: influx from $K_9$ theoretical continua.
\end{itemize}

Stability condition:
\[
\sigma(J(K_{10})) < \sigma_{\max}(K_{10}).
\]

% ================================================================
\subsubsection{Cycles $C(K_{10})$}

\[
C(K_{10})=
\{
C_{\mathrm{metaaxiom}},\;
C_{\mathrm{metarule}},\;
C_{\mathrm{metamodel}},\;
C_{\mathrm{evaluation}},\;
C_{\mathrm{coherence}},\;
C_{\mathrm{self}}
\}.
\]

Interpretation:
\begin{itemize}
    \item cycles of meta-axiom revision,
    \item cycles of meta-rule transformation,
    \item cycles of meta-model reconstruction,
    \item evaluation cycles,
    \item global coherence cycles,
    \item self-referential cycles (test–adjust–retreat–fix).
\end{itemize}

% ================================================================
\subsubsection{Time $\tau(K_{10})$}

Metatheoretical time emerges from the slowest stable meta-cycle:
\[
\tau(K_{10}) = \min_j \tau_{C_j},
\quad
\Pi(C_j) > \Theta_{\mathrm{time}}.
\]

Metatheoretical time is slower than theoretical ($K_9$) time because recursive structures change only under strong tension.

% ================================================================
\subsubsection{Continuumness $k(K_{10})$}

\[
k_{10} =
H(\Omega(K_{10}))
\cdot
\frac{|C_{\max}^{(10)}|}{|C_{\mathrm{all}}|}
\cdot
\frac{|A_{\mathrm{active}}|}{|A_{\max}|}
\cdot
\left(1-\frac{T_{10}}{\Theta_{10}}\right)_+
\cdot
\left(1-\frac{\sigma(J)}{\sigma_{\max}}\right).
\]

Interpretation:
\begin{itemize}
    \item global coherence across meta-cycles,
    \item stability of self-representation,
    \item meta-consistency levels,
    \item recursive integrity.
\end{itemize}

% ================================================================
\subsubsection{Structural Tension $T(K_{10})$}

Sources:
\begin{itemize}
    \item contradictions between meta-axioms,
    \item incompatible meta-rules,
    \item failures in recursive definitions,
    \item loss of global coherence across theories,
    \item tension between unification and plurality,
    \item paradoxes of self-reference (Russell-type, Gödel-type).
\end{itemize}

Collapse occurs when:
\[
T(K_{10}) > \Theta_{\mathrm{collapse}}.
\]

% ================================================================
\subsubsection{Energy $E(K_{10})$}

\[
E(K_{10})=
E_{\mathrm{metaaxiom}}+
E_{\mathrm{metarule}}+
E_{\mathrm{metamodel}}+
E_{\mathrm{eval}}+
E_{\mathrm{coherence}}+
E_{\mathrm{self}}+
E_{9\to10}.
\]

Energy expenditures:
\begin{itemize}
    \item maintaining meta-axioms,
    \item maintaining recursive structures,
    \item running self-evaluation cycles,
    \item enforcing global coherence,
    \item sustaining cross-theory comparisons,
    \item integrating theoretical structures from $K_9$.
\end{itemize}

% ================================================================
\subsubsection{Operators on $K_{10}$ ($\Psi$, $\Phi$, $\Lambda$, $U$, $\Chi$)}

\begin{itemize}
    \item $\Psi_{10\to11}$ — birth of $K_{11}$: structural recursion as a continuum,
    \item $\Phi$ — evolution of metatheories,
    \item $\Lambda$ — meta-stabilization (closing meta-loops),
    \item $U$ — unification of disparate $K_9$ theories,
    \item $\Chi$ — branching of metatheories (methodological pluralism vs. unity).
\end{itemize}

% ================================================================
\subsubsection{Processes on $K_{10}$}

\begin{itemize}
    \item meta-axiomatisation,
    \item reconstruction of theoretical landscapes,
    \item unification and pluralistic decomposition,
    \item recursive self-evaluation,
    \item paradigm-meta-paradigm transitions,
    \item consistency enforcement across theoretical classes.
\end{itemize}

% ================================================================
\subsubsection{Predictions for $K_{10}$}

\begin{enumerate}
    \item Stable metatheories must satisfy $\Theta_{\mathrm{metaconsistency}}$.
    \item Meta-unification is possible only when $P_{\mathrm{global}}$ is high.
    \item Self-reference stabilizes only at fixed points of $\Phi$.
    \item Recursive collapse is preceded by rising $T_{10}$.
    \item Transition to $K_{11}$ requires meta-stability and recursive definability.
\end{enumerate}

% ================================================================
\subsubsection{Experiments for $K_{10}$}

Possible proxies:
\begin{itemize}
    \item simulations of meta-rule dynamics,
    \item recursive consistency-check algorithms,
    \item global-coherence analysis of multiple theories,
    \item simulations of self-referential systems,
    \item fixed-point detection in evolving metatheories.
\end{itemize}

% ================================================================
\subsubsection{Collapse and Death of $K_{10}$}

Death when:
\[
\Omega(K_{10})=\varnothing.
\]

Mechanisms:
\begin{itemize}
    \item paradoxical self-reference,
    \item incoherent meta-rule sets,
    \item collapse of global coherence,
    \item inability to evaluate or compare theories,
    \item divergence under recursion,
    \item incompatibility with $M_{10}$.
\end{itemize}

% ================================================================
\subsubsection{Falsifiability of $K_{10}$}

To falsify:
\begin{itemize}
    \item existence of stable metatheories violating meta-consistency,
    \item recursive systems stable without fixed points,
    \item global coherence emerging below $\Theta_{\mathrm{coherence}}$,
    \item transitions to $K_{11}$ without satisfying recursive thresholds.
\end{itemize}

% ================================================================
\subsubsection{Branching / Ontological Position of $K_{10}$}

Branches:
\begin{itemize}
    \item unifying vs. pluralistic metatheories,
    \item axiomatic vs. algorithmic metatheories,
    \item static vs. dynamic recursion frameworks,
    \item single-level vs. multi-level meta-systems.
\end{itemize}

Ontological chain:
\[
K_9 \to K_{10} \to K_{11}.
\]

% ================================================================
\subsubsection{Relation to M-spaces}

$M_{10}$ defines:
\begin{itemize}
    \item allowable meta-axioms,
    \item recursion constraints,
    \item global coherence conditions,
    \item structural fixed-point conditions,
    \item bandwidth of self-reference,
    \item compatibility with $M_9$ and $M_{11}$,
    \item meta-level thresholds and universal constraints.
\end{itemize}

Compatibility with $M_{10}$ determines whether metatheoretical continua can exist.
