% ================================================================
% ==== FILE: content/k_levels/k11.tex
% ================================================================

% ==============================
%  Ontology of Continua — Core
%  K-level Module: K11 (Recursive Structural Continua)
%  FULL MODULE — FINAL
% ==============================

\section{$K_{11}$ Overview}
\label{sec:k11-overview}

$K_{11}$ is the level of \emph{recursive structural continua}: systems whose
states include entire classes of metatheories ($K_{10}$) together with
operators acting recursively on them. Unlike $K_{10}$, which evaluates,
modifies and unifies theories, $K_{11}$ organises \emph{hierarchies of
metastructures} and provides:
\begin{itemize}
    \item recursive embeddings $K_{10}\hookrightarrow K_{10}$,
    \item operators acting on families of metatheories,
    \item meta-coherence across recursive layers,
    \item fixed-point conditions for self-referential hierarchies,
    \item branching of structural possibilities (Axiom 21),
    \item the structural substrate for $K_{12}$ (limit continua).
\end{itemize}

$K_{11}$ represents the first level where the \emph{entire tower} $K_0 \to K_{10}$
can be treated as a manipulable object.

% ================================================================
\section{State Space $\Omega(K_{11})$}

\[
\Omega(K_{11})=
\{
\mathcal{M},\;
\mathcal{R},\;
\mathcal{H},\;
\mathcal{B},\;
\mathcal{F},\;
\mu^{(11)}_t
\}.
\]

Components:
\begin{itemize}
    \item $\mathcal{M}$ — sets of metatheories (structured $K_{10}$-clusters),
    \item $\mathcal{R}$ — recursive rules acting on $\mathcal{M}$,
    \item $\mathcal{H}$ — hierarchical embeddings and projections,
    \item $\mathcal{B}$ — branching structures of ontological possibilities,
    \item $\mathcal{F}$ — structural fixed-point sets,
    \item $\mu^{(11)}_t$ — dynamical measure over recursive configurations.
\end{itemize}

$K_{11}$ therefore encodes \emph{spaces of spaces} — meta-hierarchies.

% ================================================================
\section{Boundary $\partial\Omega(K_{11})$}

Boundaries arise when:
\begin{itemize}
    \item recursion becomes ill-founded (non-wellfounded chains),
    \item hierarchical embeddings contradict each other,
    \item branching leads to global inconsistency (violation of Axiom 21),
    \item fixed points fail to exist under recursion,
    \item meta-coherence cannot be preserved across levels,
    \item incompatibility with $M_{11}$ constraints appears,
    \item recursion depth or complexity exceeds the admissible bounds of $M_{11}$.
\end{itemize}

Crossing $\partial\Omega(K_{11})$ destroys the recursive continuum.

% ================================================================
\section{Axes $A(K_{11})$}

\[
A(K_{11}) =
\{
A_{\mathrm{recursive}},\;
A_{\mathrm{hier}},\;
A_{\mathrm{branch}},\;
A_{\mathrm{fix}},\;
A_{\mathrm{metaembed}},\;
A_{\mathrm{coh}}
\}.
\]

Interpretation:
\begin{itemize}
    \item $A_{\mathrm{recursive}}$ — degree of recursion,
    \item $A_{\mathrm{hier}}$ — depth and structure of hierarchies,
    \item $A_{\mathrm{branch}}$ — ontological branching axis (Axiom 21),
    \item $A_{\mathrm{fix}}$ — fixed-point formation axis,
    \item $A_{\mathrm{metaembed}}$ — meta-embedding transformations,
    \item $A_{\mathrm{coh}}$ — coherence across recursive strata.
\end{itemize}

% ================================================================
\section{Potentials $P(K_{11})$}

\[
P(K_{11})=
\{
P_{\mathrm{rec}},\;
P_{\mathrm{hier}},\;
P_{\mathrm{branch}},\;
P_{\mathrm{fix}},\;
P_{\mathrm{coh}},\;
P_{\mathrm{metaembed}}
\}.
\]

Descriptions:
\begin{itemize}
    \item recursive potential — capacity for repeated structural application,
    \item hierarchical potential — ability to support deep multilevel embeddings,
    \item branching potential — richness of ontological alternatives,
    \item fixed-point potential — ability to stabilise recursive loops,
    \item coherence potential — maintaining structural consistency across layers,
    \item embedding potential — ability to integrate or compare metatheories.
\end{itemize}

% ================================================================
\section{Thresholds $\Theta(K_{11})$}

\[
\Theta(K_{11})=
\{
\Theta_{\mathrm{rec}},\;
\Theta_{\mathrm{branch}},\;
\Theta_{\mathrm{coh}},\;
\Theta_{\mathrm{fix}},\;
\Theta_{\mathrm{hier}},\;
\Theta_{\mathrm{collapse}}
\}.
\]

Characterisation:
\begin{itemize}
    \item $\Theta_{\mathrm{rec}}$ — minimum condition for safe recursion,
    \item $\Theta_{\mathrm{branch}}$ — branching threshold (Axiom 21.4),
    \item $\Theta_{\mathrm{coh}}$ — inter-level coherence threshold,
    \item $\Theta_{\mathrm{fix}}$ — fixed-point existence threshold,
    \item $\Theta_{\mathrm{hier}}$ — minimal hierarchical stability,
    \item $\Theta_{\mathrm{collapse}}$ — limit of recursive viability.
\end{itemize}

% ================================================================
\section{Flows $J(K_{11})$}

\begin{itemize}
    \item $J_{\mathrm{rec}}$: evolution of recursive structures,
    \item $J_{\mathrm{hier}}$: flow along hierarchical embeddings,
    \item $J_{\mathrm{branch}}$: dynamics of branching topologies,
    \item $J_{\mathrm{fix}}$: formation and destruction of fixed points,
    \item $J_{\mathrm{embed}}$: embedding and projection flows,
    \item $J_{\mathrm{coh}}$: coherence-preserving flows,
    \item $J_{10\to11}$: influx of metatheoretical structures from $K_{10}$.
\end{itemize}

Stability condition:
\[
\sigma(J(K_{11})) < \sigma_{\max}(K_{11}).
\]

% ================================================================
\section{Cycles $C(K_{11})$}

\[
C(K_{11})=
\{
C_{\mathrm{rec}},\;
C_{\mathrm{hier}},\;
C_{\mathrm{branch}},\;
C_{\mathrm{fix}},\;
C_{\mathrm{embed}},\;
C_{\mathrm{coh}}
\}.
\]

Interpretation:
\begin{itemize}
    \item recursive refinement cycles,
    \item cycles of hierarchical reconstruction,
    \item branching cycles of structural differentiation,
    \item fixed-point search cycles,
    \item embedding–projection cycles,
    \item coherence restoration cycles.
\end{itemize}

% ================================================================
\section{Time $\tau(K_{11})$}

\[
\tau(K_{11})
=
\min_j \tau_{C_j},
\quad
\Pi(C_j)>\Theta_{\mathrm{time}}.
\]

Recursive time differs from metatheoretical time:
\begin{itemize}
    \item it is dominated by the slowest stable recursive process,
    \item deeper hierarchies produce slower timescales,
    \item fixed points create long-lived quasi-static plateaus.
\end{itemize}

% ================================================================
\section{Continuumness $k(K_{11})$}

\[
k_{11}=
H(\Omega(K_{11}))
\cdot
\frac{|C_{\max}^{(11)}|}{|C_{\mathrm{all}}|}
\cdot
\frac{|A_{\mathrm{active}}|}{|A_{\max}|}
\cdot
\frac{P_{\mathrm{fix}}}{P_{\mathrm{fix,max}}}
\cdot
\left(1-\frac{T_{11}}{\Theta_{11}}\right)_+.
\]

Interpretation:
\begin{itemize}
    \item recursive stability,
    \item existence of fixed points,
    \item branching without collapse,
    \item inter-level coherence.
\end{itemize}

% ================================================================
\section{Structural Tension $T(K_{11})$}

Sources:
\begin{itemize}
    \item incompatible recursive embeddings,
    \item contradictions between branches,
    \item failure of fixed-point formation,
    \item excessive branching (Axiom 21.6),
    \item collapse of meta-coherence across hierarchies.
\end{itemize}

If:
\[
T(K_{11}) > \Theta_{\mathrm{collapse}},
\]
then the recursive continuum collapses.

% ================================================================
\section{Energy $E(K_{11})$}

\[
E(K_{11})=
E_{\mathrm{rec}}+
E_{\mathrm{hier}}+
E_{\mathrm{branch}}+
E_{\mathrm{fix}}+
E_{\mathrm{coh}}+
E_{10\to11}.
\]

Energy is required for:
\begin{itemize}
    \item maintaining deep recursion,
    \item preserving branching structures,
    \item stabilising fixed points,
    \item enforcing coherence,
    \item integrating $K_{10}$ metatheories.
\end{itemize}

% ================================================================
\section{Operators on $K_{11}$ ($\Psi$, $\Phi$, $\Lambda$, $U$, $\Chi$)}

\begin{itemize}
    \item $\Psi_{11\to12}$ — formation of limit continuum $K_{12}$,
    \item $\Phi$ — evolution in recursive space,
    \item $\Lambda$ — structural closure of recursion loops,
    \item $U$ — universal embedding across recursive hierarchies,
    \item $\Chi$ — ontological branching operator (Axiom 21).
\end{itemize}

% ================================================================
\section{Processes on $K_{11}$}

\begin{itemize}
    \item recursive reconstruction of meta-structures,
    \item formation of hierarchical strata,
    \item ontological branching (controlled and uncontrolled),
    \item fixed-point search and stabilization,
    \item recursive integration of $K_{10}$ structures,
    \item cross-level coherence enforcement.
\end{itemize}

% ================================================================
\section{Predictions for $K_{11}$}

\begin{enumerate}
    \item Stable recursion requires $P_{\mathrm{fix}}>0$ and $\Theta_{\mathrm{fix}}$ satisfied.
    \item Excessive branching leads to $T(K_{11})\uparrow$ and eventual collapse.
    \item Deep hierarchies produce longer timescales and slow dynamics.
    \item Transition to $K_{12}$ requires existence of a global recursive fixed point.
    \item Inter-level incoherence predicts collapse before $K_{12}$ formation.
\end{enumerate}

% ================================================================
\section{Experiments for $K_{11}$}

Possible proxies:
\begin{itemize}
    \item simulations of recursive operators,
    \item fixed-point detection algorithms for recursive maps,
    \item agent-based recursive hierarchy models,
    \item branching-process simulations,
    \item meta-meta-consistency solvers.
\end{itemize}

% ================================================================
\section{Collapse and Death of $K_{11}$}

Death when:
\[
\Omega(K_{11})=\varnothing.
\]

Mechanisms:
\begin{itemize}
    \item recursive divergence,
    \item loss of fixed points,
    \item catastrophic branching,
    \item collapse of hierarchical coherence,
    \item incompatibility with $M_{11}$.
\end{itemize}

% ================================================================
\section{Falsifiability of $K_{11}$}

The level is falsified if:
\begin{itemize}
    \item stable recursive hierarchies exist without fixed points,
    \item branching occurs below $\Theta_{\mathrm{branch}}$,
    \item cross-level coherence arises without meeting $\Theta_{\mathrm{coh}}$,
    \item transitions to $K_{12}$ occur without recursive stability,
    \item recursion operates with well-foundedness violations that do not collapse.
\end{itemize}

% ================================================================
\section{Branching / Ontological Position of $K_{11}$}

Branches:
\begin{itemize}
    \item shallow vs.\ deep recursion regimes,
    \item narrow vs.\ wide branching structures,
    \item coherent vs.\ incoherent hierarchical embeddings,
    \item fixed-point-rich vs.\ fixed-point-poor regimes.
\end{itemize}

Position in ontological chain:
\[
K_{10} \to K_{11} \to K_{12}.
\]

% ================================================================
\section{Relation to M-spaces}

$M_{11}$ specifies:
\begin{itemize}
    \item recursion admissibility,
    \item global coherence bounds,
    \item branching limits,
    \item hierarchical embedding capacities,
    \item fixed-point existence constraints,
    \item compatibility with $M_{10}$ and $M_{12}$.
\end{itemize}

A continuum $K_{11}$ exists iff its structures satisfy the admissibility region of $M_{11}$.
