% ================================================================
% ==== FILE: content/k_levels/k12.tex
% ================================================================

% ==============================
%  Ontology of Continua — Core
%  K-level Module: K12 (Limit Continua)
%  FULL MODULE — FINAL
% ==============================

\section{$K_{12}$ Overview}
\label{sec:k12-overview}

$K_{12}$ is the \emph{limit continuum}: the structural fixed point of the
recursive meta-hierarchy developed in $K_{11}$.  
It is the first level where:

\begin{itemize}
    \item recursive meta-structures stabilise,
    \item branching (Axiom~21) reaches closure,
    \item a global structural fixed point exists,
    \item all operators $(\Psi,\Phi,\Lambda,U,\Chi)$ act as symmetries,
    \item the entire ladder $K_0\dots K_{11}$ becomes representable inside one
          invariant continuum.
\end{itemize}

$K_{12}$ is therefore not merely “the next level”, but the structural limit of
the K-tower in Core v2.6.

% ================================================================
\section{State Space $\Omega(K_{12})$}

\[
\Omega(K_{12}) =
\{
\Omega^*,\;
A^*,\;
P^*,\;
\Theta^*,\;
J^*,\;
C^*,\;
\mu^{(12)}_t
\},
\]

where:
\begin{itemize}
    \item $\Omega^*$ is the limit state-space containing all admissible embeddings
          of $K_0\!\to\!K_{11}$,
    \item $A^*$ is the global axis set induced by the closure of all Axes
          from previous levels,
    \item $P^*$ are potentials that survive recursive convergence,
    \item $\Theta^*$ are stable thresholds preserved under recursion,
    \item $J^*$ are flows that remain invariant under all structural operators,
    \item $C^*$ is the fixed family of cycles that defines the time of $K_{12}$,
    \item $\mu^{(12)}_t$ is the measure over limit configurations.
\end{itemize}

All components satisfy global invariance:
\[
X^* = \mathcal{R}(X^*),\quad \forall X \in \{\Omega, A, P, \Theta, J, C\}.
\]

% ================================================================
\section{Boundary $\partial\Omega(K_{12})$}

The boundary consists of:
\begin{itemize}
    \item configurations where recursion does not converge,
    \item points where branching produces incompatible global topologies,
    \item states violating limit invariance under operators,
    \item divergence of fixed-point cycles,
    \item non-admissibility in $M_{12}$.
\end{itemize}

Crossing $\partial\Omega(K_{12})$ destroys the limit structure and reverts the
system to a failing $K_{11}$-regime.

% ================================================================
\section{Axes $A(K_{12})$}

\[
A(K_{12}) =
\overline{\bigcup_{i=0}^{11} A(K_i)}
\quad \text{subject to recursive closure}.
\]

This axis set is:
\begin{itemize}
    \item closed under recursion from $K_{11}$,
    \item stable under all operators $(\Psi,\Phi,\Lambda,U,\Chi)$,
    \item minimal among axis sets satisfying global invariance.
\end{itemize}

Each axis becomes a \emph{symmetry direction} in $K_{12}$.

% ================================================================
\section{Potentials $P(K_{12})$}

\[
P(K_{12}) = \lim_{n\to\infty} P^{(n)}(K_{11}),
\]

i.e.\ the potentials that remain finite, coherent and closed under recursion:
\begin{itemize}
    \item $P_{\mathrm{fix}}^*$ — fixed-point-supporting potential,
    \item $P_{\mathrm{coh}}^*$ — global coherence potential,
    \item $P_{\mathrm{branch}}^*$ — stabilised branching potential,
    \item $P_{\mathrm{embed}}^*$ — universal embedding potential.
\end{itemize}

% ================================================================
\section{Thresholds $\Theta(K_{12})$}

\[
\Theta(K_{12}) =
\lim_{n\to\infty} \Theta^{(n)}(K_{11}),
\]

giving:
\begin{itemize}
    \item $\Theta^*_{\mathrm{fix}}$ — fixed-point existence threshold,
    \item $\Theta^*_{\mathrm{coh}}$ — global coherence threshold,
    \item $\Theta^*_{\mathrm{branch}}$ — branching-closure threshold,
    \item $\Theta^*_{\mathrm{inv}}$ — invariance-threshold for operators,
    \item $\Theta^*_{\mathrm{collapse}}$ — structural collapse limit.
\end{itemize}

All thresholds become \emph{globally stable constants} for the limit continuum.

% ================================================================
\section{Flows $J(K_{12})$}

\[
J(K_{12}) = \{J^*\mid \Phi(J^*)=J^*\}.
\]

Thus every flow must be:
\begin{itemize}
    \item invariant under recursive evolution,
    \item compatible with limit thresholds,
    \item globally coherent across projections and embeddings,
    \item fixed under operator action.
\end{itemize}

This is the only level where \emph{all flows are symmetry flows}.

% ================================================================
\section{Cycles $C(K_{12})$}

\[
C(K_{12}) = \{ C^*\mid \Pi(C^*)>\Theta_{\mathrm{time}}^*,\;
\Lambda(C^*)=C^*\}.
\]

These are the cycles that:
\begin{itemize}
    \item persist through all recursive refinements,
    \item define the emergent time of $K_{12}$,
    \item remain invariant under branching and operator action,
    \item serve as the structural backbone of the continuum.
\end{itemize}

% ================================================================
\section{Time $\tau(K_{12})$}

Time is determined by invariant cycles:
\[
\tau(K_{12}) = \min_j \tau_{C_j^*}.
\]

Properties:
\begin{itemize}
    \item all temporal directions are symmetries,
    \item no new time directions can arise,
    \item time is globally coherent and invariant under recursion.
\end{itemize}

% ================================================================
\section{Continuumness $k(K_{12})$}

\[
k_{12} =
H(\Omega(K_{12}))
\cdot
\frac{|C^*|}{|C_{\max}|}
\cdot
\frac{|A^*|}{|A_{\max}|}
\cdot
\frac{P_{\mathrm{fix}}^*}{P_{\mathrm{fix,max}}}
\cdot
\left(1-\frac{T_{12}}{\Theta^*_{\mathrm{collapse}}}\right)_+.
\]

$k_{12}$ attains the maximum possible value among all $K$-levels:
\[
k_{12} = \max_{i=0\dots12} k(K_i).
\]

% ================================================================
\section{Structural Tension $T(K_{12})$}

Sources:
\begin{itemize}
    \item unresolved branching incompatibilities,
    \item operator-invariance failures,
    \item instability of fixed points (rare but possible),
    \item projection–embedding inconsistencies,
    \item violations of $M_{12}$ bounds.
\end{itemize}

Collapse occurs if:
\[
T(K_{12}) > \Theta^*_{\mathrm{collapse}}.
\]

% ================================================================
\section{Energy $E(K_{12})$}

\[
E(K_{12}) =
E_{\mathrm{fix}}^*
+E_{\mathrm{coh}}^*
+E_{\mathrm{branch}}^*
+E_{\mathrm{embed}}^*
+E_{\mathrm{sym}}.
\]

Interpretation:
\begin{itemize}
    \item minimal energy required to preserve fixed-point structure,
    \item energy of global coherence maintenance,
    \item energy stabilising branching closure,
    \item energy supporting universal embeddings,
    \item symmetry energy of operator invariance.
\end{itemize}

$K_{12}$ is the most energy-stable admissible continuum.

% ================================================================
\section{Operators on $K_{12}$ ($\Psi$, $\Phi$, $\Lambda$, $U$, $\Chi$)}

At the limit level:

\begin{itemize}
    \item $\Psi$ — becomes identity on admissible states,
    \item $\Phi$ — becomes the global invariance flow,
    \item $\Lambda$ — closure of all limit cycles,
    \item $U$ — universal embedding operator,
    \item $\Chi$ — stabilised branching operator.
\end{itemize}

Operator algebra becomes Abelian up to equivalence classes:
\[
[\Psi,\Phi]=[\Phi,\Lambda]=\dots = 0.
\]

% ================================================================
\section{Processes on $K_{12}$}

\begin{itemize}
    \item convergence of recursive structures,
    \item closure of branching topologies,
    \item formation of global fixed points,
    \item collapse of non-admissible hierarchies,
    \item emergence of full symmetry under operators,
    \item stabilisation of time and cycle structure.
\end{itemize}

% ================================================================
\section{Predictions for $K_{12}$}

\begin{enumerate}
    \item Any admissible recursive meta-hierarchy eventually converges to
          structures representable in $K_{12}$.
    \item No new axes or potentials can emerge beyond this level.
    \item All operator actions reduce to global symmetries.
    \item Branching becomes topologically closed; no new branches arise.
    \item Failure to achieve a global fixed point forces collapse back to $K_{11}$.
\end{enumerate}

% ================================================================
\section{Experiments for $K_{12}$}

Only indirect experiments exist:
\begin{itemize}
    \item simulations of recursive convergence,
    \item fixed-point computation in high-order meta-spaces,
    \item invariance tests under operator algebra,
    \item stability tests for branching closures,
    \item modelling admissibility under $M_{12}$.
\end{itemize}

% ================================================================
\section{Collapse and Death of $K_{12}$}

Death if:
\[
\Omega(K_{12})=\varnothing.
\]

Mechanisms:
\begin{itemize}
    \item recursive divergence at deep levels,
    \item operator-invariance violations,
    \item failure of branching closure,
    \item global incoherence of fixed points,
    \item incompatibility with the meta-space $M_{12}$.
\end{itemize}

% ================================================================
\section{Falsifiability of $K_{12}$}

Falsified if:
\begin{itemize}
    \item a meta-hierarchy converges without forming invariant fixed points,
    \item branching remains unclosed yet stability is preserved,
    \item new axes or potentials emerge beyond closure,
    \item operator algebra fails to become symmetric,
    \item admissible recursion produces non-convergent hierarchies.
\end{itemize}

% ================================================================
\section{Branching / Ontological Position of $K_{12}$}

Branching structure:
\begin{itemize}
    \item all branches close,
    \item no new branches form,
    \item the topology becomes globally compact and invariant.
\end{itemize}

Ontological position:
\[
K_{11} \;\longrightarrow\; K_{12} \quad \text{(limit continuum)}.
\]

Beyond $K_{12}$ the theory does not posit new levels; the limit continuum
contains all admissible structural possibilities.

% ================================================================
\section{Relation to M-spaces}

$M_{12}$ defines:
\begin{itemize}
    \item full admissibility region for limit structures,
    \item invariance requirements for operators,
    \item global coherence constraints,
    \item branching closure bounds,
    \item fixed-point existence conditions,
    \item compatibility with all $M_0\dots M_{11}$.
\end{itemize}

A continuum exists at level $K_{12}$ iff its entire structure lies within the
admissible region of $M_{12}$.
