% ================================================================
% ==== FILE: content/jets/jets_k8.tex
% ================================================================

\subsubsection{Jets on \texorpdfstring{$K_8$}{K_8}}
\label{sec:jets-k8}

The level $K_8$ corresponds to civilisational continua: large-scale
aggregates of agents, institutions, symbolic structures, technologies,
and energy flows. Jets on $K_8$ allow us to analyse the infinitesimal
evolution of infrastructures, symbolic coherence, systemic fragility,
and long-term stability of complex societies.


% ================================================================
\subsubsection{State Space and Jet Coordinates of \texorpdfstring{$K_8$}{K_8}}
% ================================================================

A civilisational state $\omega_8$ is represented as
\[
\omega_8 = 
(I,\ T,\ S,\ P,\ J),
\]
where
\begin{itemize}
    \item $I$ — infrastructure configurations (transport, energy, logistics),
    \item $T$ — technological systems and production networks,
    \item $S$ — symbolic structures (language, media, cultural codes),
    \item $P$ — macro-potentials (energy capacity, institutional tension,
                symbolic coherence, systemic stress),
    \item $J$ — flows (information, energy, materials, finance).
\end{itemize}

Jet coordinates:
\[
J^{m,n}(K_8)=
\{\partial_t^k \partial_{A_8}^\ell (I,T,S,P,J)\},
\quad
0\le k\le m,\;0\le\ell\le n,
\]
where $A_8$ includes infrastructural, technological, symbolic,
institutional, and macro-energetic axes.


% ================================================================
\subsubsection{Jets of Infrastructure \texorpdfstring{$I$}{I}}
% ================================================================

Infrastructure dynamics:
\[
\dot{I} = F_I(I, T, P_{\mathrm{energy}}, J_{\mathrm{material}}).
\]

Jet:
\[
J^1(I)=\{\dot{I},\ \partial_{A_8}I\}.
\]

Interpretation:
\begin{itemize}
    \item robustness and redundancy of networks,
    \item sensitivity to energy constraints,
    \item propagation of failures (cascades),
    \item infrastructure-driven phase transitions.
\end{itemize}


% ================================================================
\subsubsection{Jets of Technological Systems \texorpdfstring{$T$}{T}}
% ================================================================

Technological systems evolve via:
\[
\dot{T} = F_T(T, I, S, P_{\mathrm{innovation}}, J_{\mathrm{info}}).
\]

Jet:
\[
\partial_{A_8}T=
\partial_I T\,\partial_{A_8}I
+
\partial_S T\,\partial_{A_8}S
+
\partial_P T\,\partial_{A_8}P.
\]

This captures:
\begin{itemize}
    \item innovation dynamics,
    \item technological lock-in,
    \item dependence on symbolic structures (education, language),
    \item transitions between technological regimes.
\end{itemize}


% ================================================================
\subsubsection{Jets of Symbolic Structures \texorpdfstring{$S$}{S}}
% ================================================================

Symbolic structures include shared languages, media, narratives and
cultural systems:
\[
\dot{S}=F_S(S, T, P_{\mathrm{coherence}}, J_{\mathrm{info}}).
\]

Jet:
\[
J^1(S)=\{\dot{S},\ \partial_{A_8}S\}.
\]

Interpretation:
\begin{itemize}
    \item symbolic coherence and fragmentation,
    \item emergence of shared narratives,
    \item media-driven instability,
    \item cultural drift and codification.
\end{itemize}


% ================================================================
\subsubsection{Jets of Civilisational Potentials \texorpdfstring{$P_8$}{P_8}}
% ================================================================

Macro-potentials include:
\[
P_8 = 
(P_{\mathrm{energy}},\ 
 P_{\mathrm{coherence}},\ 
 P_{\mathrm{inst}},\ 
 P_{\mathrm{stress}},\ 
 P_{\mathrm{innovation}}).
\]

Jets:
\[
J^1(P_8)=\{\dot{P}_8,\ \partial_{A_8}P_8\}.
\]

Potentials encode:
\begin{itemize}
    \item available energy reserves and conversion efficiency,
    \item symbolic cohesion,
    \item institutional tension and legitimacy,
    \item systemic stress accumulation,
    \item capacity for technological advance.
\end{itemize}


% ================================================================
\subsubsection{Jets of Civilisational Flows \texorpdfstring{$J_8$}{J_8}}
% ================================================================

Flows contain:
\[
J_8=
(J_{\mathrm{energy}},\ 
 J_{\mathrm{material}},\ 
 J_{\mathrm{info}},\ 
 J_{\mathrm{finance}},\ 
 J_{\mathrm{institution}}).
\]

Dynamics:
\[
\dot{J}_8 = F_J(J_8, I, T, S, P_8).
\]

Jet:
\[
J^1(J_8)=\{\dot{J}_8,\ \partial_{A_8}J_8\}.
\]

Interpretation:
\begin{itemize}
    \item global coupling of subsystems,
    \item congestion and diffusion regimes,
    \item propagation of shocks,
    \item formation of large-scale attractors.
\end{itemize}


% ================================================================
\subsubsection{Jets of Thresholds \texorpdfstring{$\Theta_8$}{\Theta_8}}
% ================================================================

Key civilisational thresholds:
\[
\Theta_8=
\{
\Theta_{\mathrm{energy}},\,
\Theta_{\mathrm{infrastructure}},\,
\Theta_{\mathrm{coherence}},\,
\Theta_{\mathrm{inst}},\,
\Theta_{\mathrm{fragility}}
\}.
\]

Jets:
\[
J^1(\Theta_8)=\{\dot{\Theta}_8,\ \partial_{A_8}\Theta_8\}.
\]

Interpretation:
\begin{itemize}
    \item proximity to infrastructural collapse,
    \item symbolic fragmentation boundary,
    \item institutional instability,
    \item energy constraints on growth,
    \item systemic fragility amplification.
\end{itemize}


% ================================================================
\subsubsection{Jets of Civilisational Cycles \texorpdfstring{$C_8$}{C_8}}
% ================================================================

Typical cycles of $K_8$:
\[
C_{\mathrm{infra}},\;
C_{\mathrm{energy}},\;
C_{\mathrm{symbolic}},\;
C_{\mathrm{tech}},\;
C_{\mathrm{inst}},\;
C_{\mathrm{stress}}.
\]

A jet of a cycle $C_j$:
\[
J^1(C_j)=
\{\partial_t X_j,\ \partial_{A_8}X_j\},
\qquad 
X_j\in\{I,T,S,P,J\}.
\]

Cycle jets reveal:
\begin{itemize}
    \item long-term civilisational rhythms,
    \item shocks and recovery processes,
    \item positive feedback loops,
    \item transitions from expansion to stagnation,
    \item systemic crises (collapse attractors),
    \item pathways to new civilisational phases.
\end{itemize}


% ================================================================
\subsubsection{Jets and Boundary Stability \texorpdfstring{$\partial\Omega(K_8)$}{\partial\Omega(K_8)}}
% ================================================================

Instability occurs when:
\[
\partial_{A_8}(P_{\mathrm{energy}} - \Theta_{\mathrm{energy}})<0,
\]
or
\[
\partial_{A_8}(P_{\mathrm{coherence}} - \Theta_{\mathrm{coherence}})<0,
\]
or
\[
\partial_{A_8}(P_{\mathrm{inst}} - \Theta_{\mathrm{inst}})<0.
\]

These inequalities indicate:
\begin{itemize}
    \item resource exhaustion,
    \item symbolic fragmentation,
    \item institutional collapse.
\end{itemize}

Jets quantify the rates and directions by which the system approaches
$\partial\Omega(K_8)$.


% ================================================================
\subsubsection{Jets and the Transition \texorpdfstring{$K_7 \to K_8$}{K_7 \to K_8}}
% ================================================================

The rise of the civilisational continuum corresponds to:
\[
\partial_{A_8}S
\propto
\partial_{A_7}C_{\mathrm{coord}},
\]
\[
\partial_{A_8}I
\propto
\partial_{A_7}J_{\mathrm{resource}},
\]
\[
\partial_{A_8}P_{\mathrm{coherence}}
\propto
\partial_{A_7}P_{\mathrm{trust}}.
\]

Thus stable symbolic structures at $K_7$ become enlarged and codified
into long-range infrastructures, institutional systems, and
technological networks characteristic of $K_8$.


% ================================================================
\subsubsection{Summary}
% ================================================================

Jets on $K_8$ formalise the infinitesimal dynamics of civilisational
continua:
\begin{itemize}
    \item infrastructure robustness and propagation of failures,
    \item technological regimes and innovation dynamics,
    \item symbolic coherence and cultural drift,
    \item macro-potentials governing stability and growth,
    \item large-scale flows across energy, information and materials,
    \item civilisational thresholds and systemic fragility,
    \item long-term cycles and collapse dynamics,
    \item emergence from $K_7$ social coordination.
\end{itemize}

They provide the analytical basis for transitions to higher-level
meta-structural continua $K_9$ and $K_{10}$.
