% ================================================================
% ==== FILE: content/jets/jets_k0.tex
% ================================================================

\section{Jets on \texorpdfstring{$K_0$}{K_0}}
\label{sec:jets-k0}

The level $K_0$ represents the degenerate, pre-geometric continuum of the
Ontology of Continua (OC). It contains no axes, no potentials,
no thresholds beyond the minimal existence threshold $\Theta_0=\varepsilon>0$,
and no temporal evolution. Consequently, the jet structure on $K_0$ is
both trivial and foundational: it provides the limiting case against
which all higher-order jets are defined.

\medskip

Formally, the structural content of $K_0$ is given by:
\[
K_0 = (S, \Delta, \mathcal{C}),
\]
where $S$ is an undifferentiated structural substrate,
$\Delta$ is the minimal distinction required for the existence of a
continuum, and $\mathcal{C}$ denotes the admissible compositions of
distinctions. No metric, topological, temporal, or dynamical structure
exists at this level.

Thus $K_0$ admits only the \emph{zero jet}, representing the fact that
no derivatives of any structural quantity can be defined.



% ================================================================
\subsection{Definition of Jets on \texorpdfstring{\texorpdfstring{$K_0$}{K_0}}{K0}}
% ================================================================

Since $K_0$ possesses no time parameter, no axes, and no dynamical fields,
the general definition of $n$-jets collapses to the trivial object:
\[
J^n(K_0) = \{0\}
\qquad \text{for all } n \ge 0.
\]

This means:

\begin{itemize}
    \item no quantities can vary with time (time does not yet exist),
    \item no gradients can be defined (no axes exist),
    \item no flows can be defined (no admissible transitions),
    \item no thresholds evolve (only the existence threshold $\Theta_0$),
    \item $\partial\Omega(K_0)$ is static and degenerate.
\end{itemize}

The jet formalism therefore records the essential fact that $K_0$
exhibits \emph{zero differential structure}. All higher jet structures
$J^n(K_x)$ for $x\ge1$ must reduce to $J^n(K_0)$ in the appropriate
singular limit.



% ================================================================
\subsection{Jet Constraints and Admissibility}
% ================================================================

The existence of jets at $K_0$ reduces to the requirement that the minimal
threshold of existence is respected:
\[
T_0 < \Theta_0 = \varepsilon.
\]

If this condition fails, the continuum does not exist at all:
\[
K_0 = \varnothing.
\]

Since $K_0$ has no time evolution, the inequality is structural rather
than dynamical; it represents the minimal consistency condition of the
pre-continuum. Any attempt to introduce derivative-like behaviour would
violate admissibility, since no such behaviour is defined.



% ================================================================
\subsection{Transition to \texorpdfstring{\texorpdfstring{$K_1$}{K_1}}{K1}: Jets and the Birth of Time}
% ================================================================

The transition $K_0 \to K_1$ is governed by the operator $\Psi_{0\to1}$,
which introduces:

\begin{itemize}
    \item the first axis $A_1$,
    \item a topological substrate $(X,\tau)$,
    \item classical admissible configurations $\Omega_{\mathrm{cl}}(K_1)$,
    \item the first boundary structure $\partial\Omega_{\mathrm{cl}}$,
    \item the notion of time as the parameter of evolution.
\end{itemize}

In terms of jets, this transition is described as:

\begin{equation}
J^0(K_0) = \{0\}
\quad \longrightarrow \quad
J^1(K_1) = 
\left\{
\frac{dA_1}{dt},
\frac{d}{dt}\partial\Omega_{\mathrm{cl}},
\frac{dE}{dt},
\frac{dT_1}{dt},
\ldots
\right\}.
\end{equation}

Thus, the birth of $K_1$ is the birth of nontrivial jet structure.  
All higher continua inherit and then enrich this basic infinitesimal
geometry.

The jet structure therefore provides a rigorous mathematical
interpretation of the emergence of time and geometry from the degenerate
pre-continuum $K_0$.



% ================================================================
\subsection{Summary}
% ================================================================

Jets on $K_0$ are trivial but foundational:

\begin{itemize}
    \item all jets reduce to the zero jet,
    \item no derivatives exist,
    \item no flows, gradients, or potentials exist,
    \item the only admissibility condition is structural existence,
    \item the first nontrivial jet appears only at $K_1$.
\end{itemize}

These properties ensure that $K_0$ acts as the unique jet-free anchor of
the entire continuum hierarchy, providing the reference point for the
emergence of infinitesimal structure at every higher level.
