% ================================================================
% ==== FILE: content/jets/jets_k11.tex
% ================================================================

\section{Jets on $K_{11}$}
\label{sec:jets-k11}

The level $K_{11}$ represents fully formalised ontological systems:
closed logical structures equipped with admissible rules of inference,
operators on differences of arbitrary order, and a rigorously defined
boundary of coherent ontology. Jets on $K_{11}$ describe infinitesimal
changes in these formal systems and capture the onset of instability,
collapse or restructuring of the deepest representational layer of a
continuum.


% ================================================================
\subsection{State Space and Jet Coordinates of $K_{11}$}
% ================================================================

A state of $K_{11}$ is a formal ontological structure:
\[
\omega_{11}
=
(\mathcal{O},\ \mathcal{R},\ \Delta,\ \Pi),
\]
where:
\begin{itemize}
    \item $\mathcal{O}$ — the set of admissible ontological objects;
    \item $\mathcal{R}$ — formal relations and inference rules;
    \item $\Delta$ — higher-order difference operators
                     (beyond $\mathfrak{D}$ of $K_{10}$);
    \item $\Pi$ — admissible proof, consistency and closure schemes.
\end{itemize}

Jets on $K_{11}$ are defined as:
\[
J^{m,n}(K_{11}) =
\left\{
\partial_t^k\partial_{A_{11}}^\ell
(\mathcal{O},\mathcal{R},\Delta,\Pi)
\right\},
\quad
0\le k\le m,\; 0\le\ell\le n,
\]
where $A_{11}$ is the axis of formal-ontological variation.


% ================================================================
\subsection{Jets of Ontological Objects $\mathcal{O}$}
% ================================================================

Dynamics:
\[
\dot{\mathcal{O}}
=
F_{\mathcal{O}}(\mathcal{O},\mathcal{R},\Delta,\Pi).
\]

Jet:
\[
J^1(\mathcal{O})=\{\dot{\mathcal{O}},\ \partial_{A_{11}}\mathcal{O}\}.
\]

Interpretation:
\begin{itemize}
    \item refinement or restriction of admissible objects;
    \item birth of new permissible ontological entities;
    \item contraction of $\mathcal{O}$ as boundaries approach collapse;
    \item elimination of objects violating consistency schemes.
\end{itemize}


% ================================================================
\subsection{Jets of Ontological Relations $\mathcal{R}$}
% ================================================================

Relations and inference rules evolve as constraints from
$\Delta$ and $\Pi$ shift.

Dynamics:
\[
\dot{\mathcal{R}}
=
F_{\mathcal{R}}(\mathcal{R},\mathcal{O},\Delta,\Pi).
\]

Jet:
\[
\partial_{A_{11}}\mathcal{R}
=
\partial_{\mathcal{O}}\mathcal{R}\,\partial_{A_{11}}\mathcal{O}
+
\partial_{\Delta}\mathcal{R}\,\partial_{A_{11}}\Delta
+
\partial_{\Pi}\mathcal{R}\,\partial_{A_{11}}\Pi.
\]

Interpretation:
\begin{itemize}
    \item modifications of admissible inference patterns,
    \item structural shifts in relation schemas,
    \item axiomatic transformations,
    \item collapse of relations under inconsistency.
\end{itemize}


% ================================================================
\subsection{Jets of Higher-Order Difference Operators $\Delta$}
% ================================================================

In $K_{11}$, difference operators become fully formal:
they act on distinctions of arbitrary order.

Dynamics:
\[
\dot{\Delta}
=
F_{\Delta}(\Delta,\mathcal{O},\mathcal{R},\Pi).
\]

Jet:
\[
J^1(\Delta)=\{\dot{\Delta},\ \partial_{A_{11}}\Delta\}.
\]

Interpretation:
\begin{itemize}
    \item emergence of new orders of distinction,
    \item restructuring of difference hierarchies,
    \item incompatibility cascades leading to collapse,
    \item sharpening of ontological axes.
\end{itemize}


% ================================================================
\subsection{Jets of Consistency and Closure Schemes $\Pi$}
% ================================================================

$\Pi$ encodes formal admissibility, proof systems, soundness,
closure under inference and anti-paradox constraints.

Dynamics:
\[
\dot{\Pi}
=
F_{\Pi}(\Pi,\mathcal{O},\mathcal{R},\Delta).
\]

Jet:
\[
J^1(\Pi)=\{\dot{\Pi},\ \partial_{A_{11}}\Pi\}.
\]

Interpretation:
\begin{itemize}
    \item strengthening or weakening of consistency boundaries,
    \item refinement of admissible derivational schemes,
    \item meta-logical phase transitions,
    \item approach to inconsistency collapse.
\end{itemize}


% ================================================================
\subsection{Jets of Thresholds $\Theta_{11}$}
% ================================================================

$K_{11}$ inherits and amplifies the meta-theoretical thresholds:
\[
\Theta_{11}
=
(\Theta_{\mathrm{consist}},\ 
 \Theta_{\mathrm{closure}},\ 
 \Theta_{\mathrm{order}},\ 
 \Theta_{\mathrm{reflex}}),
\]
with interpretations:
\begin{itemize}
    \item $\Theta_{\mathrm{consist}}$ — limit of consistency load;
    \item $\Theta_{\mathrm{closure}}$ — maximal admissible closure depth;
    \item $\Theta_{\mathrm{order}}$ — boundary on order of permissible distinctions;
    \item $\Theta_{\mathrm{reflex}}$ — limit of self-referential admissibility.
\end{itemize}

Jet:
\[
J^1(\Theta_{11})
=
\{\dot{\Theta}_{11},\ \partial_{A_{11}}\Theta_{11}\}.
\]

Instability signals:
\[
T_{11}
=
w_c (C_{\mathrm{consist}}-\Theta_{\mathrm{consist}})
+
w_o (O_{\mathrm{order}}-\Theta_{\mathrm{order}})
+
w_r (R_{\mathrm{reflex}}-\Theta_{\mathrm{reflex}}),
\]
with $T_{11}>\Theta_{\mathrm{death}}$ implying
$\Omega(K_{11})=\varnothing$.


% ================================================================
\subsection{Jets of Formal Flows $J^{(11)}$}
% ================================================================

Flows:
\[
J^{(11)}
=
\left(
\frac{d\mathcal{O}}{dt},\ 
\frac{d\mathcal{R}}{dt},\ 
\frac{d\Delta}{dt},\ 
\frac{d\Pi}{dt}
\right).
\]

Jet:
\[
J^1(J^{(11)})
=
\big\{
\partial_t^2(\mathcal{O},\mathcal{R},\Delta,\Pi),\ 
\partial_{A_{11}}\partial_t(\mathcal{O},\mathcal{R},\Delta,\Pi)
\big\}.
\]

Interpretation:
\begin{itemize}
    \item acceleration of formal structural change,
    \item divergence of inference schemes,
    \item formation of inconsistency fronts,
    \item transition to global ontological collapse.
\end{itemize}


% ================================================================
\subsection{Jets of Formal Cycles $C_{11}$}
% ================================================================

Cycles appear only in highly stable $K_{11}$ regimes:
\[
C_j:\quad
\text{ontology}
\to\text{inference}
\to\text{difference hierarchy}
\to\text{closure}
\to\text{ontology}.
\]

Jet of cycle:
\[
J^1(C_j)
=
\{\partial_t X_j,\ \partial_{A_{11}}X_j\},
\quad X_j\in
\{\mathcal{O},\mathcal{R},\Delta,\Pi\}.
\]

Interpretation:
\begin{itemize}
    \item existence of stable logical-ontological equilibrium,
    \item capacity of the system to regenerate formal coherence,
    \item sensitivity to perturbations in higher-order distinctions,
    \item precursor signals of meta-logical collapse.
\end{itemize}


% ================================================================
\subsection{Jets and Boundary Geometry $\partial\Omega(K_{11})$}
% ================================================================

Approach to the boundary occurs when:
\[
\partial_{A_{11}}
(C_{\mathrm{consist}}-\Theta_{\mathrm{consist}})>0,
\quad
\partial_{A_{11}}
(O_{\mathrm{order}}-\Theta_{\mathrm{order}})>0,
\]
\[
\partial_{A_{11}}
(R_{\mathrm{reflex}}-\Theta_{\mathrm{reflex}})>0.
\]

Interpretation:
\begin{itemize}
    \item rise of super-critical self-referential depth,
    \item runaway escalation of distinction orders,
    \item destabilisation of proof/closure systems,
    \item collapse of formal ontological space.
\end{itemize}


% ================================================================
\subsection{Relation to $K_{10} \to K_{11}$ Transition}
% ================================================================

Transition is triggered when:
\begin{itemize}
    \item meta-theoretical constructs become strictly formalised,
    \item modal operators acquire consistency-sensitive interpretation,
    \item functorial and category-theoretic structures become axiomatic,
    \item higher-order differences become admissible inference operators.
\end{itemize}

The jet relation:
\[
\partial_{A_{11}}(\mathcal{O},\mathcal{R},\Delta,\Pi)
\propto
\partial_{A_{10}}(\mathcal{F},\mathcal{C},\Lambda,\mathfrak{D})
\]
marks the embedding of meta-theory into fully formal ontology.


% ================================================================
\subsection{Summary}
% ================================================================

Jets on $K_{11}$ describe:
\begin{itemize}
    \item infinitesimal changes in ontological objects, relations and
          proof systems;
    \item restructuring of difference hierarchies and admissible distinctions;
    \item behaviour of formal thresholds and reflexive limits;
    \item dynamics of formal cycles and precursors of collapse;
    \item transition from meta-theoretical continua to
          fully formal ontological continua.
\end{itemize}

They provide a differential machinery for analysing the birth,
instability and collapse of the highest-order representational structures.
