% ================================================================
% ==== FILE: content/jets/jets_k7.tex
% ================================================================

\subsubsection{Jets on \texorpdfstring{$K_7$}{K_7}}
\label{sec:jets-k7}

The level $K_7$ introduces social continua: multi-agent structures with
collective states, norms, roles, distributed information, and
institutional constraints. Jets on $K_7$ provide the infinitesimal
analysis of social potentials, flows, coordination mechanisms, and
instability thresholds governing cooperation, conflict, and institutional
persistence.


% ================================================================
\subsubsection{State Space and Jet Coordinates of \texorpdfstring{$K_7$}{K_7}}
% ================================================================

A social-state $\omega_7$ consists of
\[
\omega_7 = 
(G,\ R,\ S,\ N,\ P,\ J),
\]
where:
\begin{itemize}
    \item $G$ — group-structure configuration (membership, topology),
    \item $R$ — role assignments and behavioural repertoires,
    \item $S$ — status and authority distributions,
    \item $N$ — norms and institutional constraints,
    \item $P$ — social potentials (trust, cohesion, sanction pressure),
    \item $J$ — social flows (information, resources, influence).
\end{itemize}

The $m$-$n$ jet bundle of $K_7$:
\[
J^{m,n}(K_7)
=
\big\{
\partial_t^k \partial_{A_7}^\ell (G,R,S,N,P,J)
\big\},
\quad
0\le k\le m,\; 0\le\ell\le n,
\]
where $A_7$ denotes the set of social axes (coordination, imitation,
status-gradient, normative force, institutional rigidity, network
connectivity).


% ================================================================
\subsubsection{Jets of Group Structure \texorpdfstring{$G$}{G}}
% ================================================================

Group structure evolves by:
\[
\dot{G} = F_G(G,S,P,J_{\mathrm{comm}}).
\]

Jets:
\[
J^1(G) = \{\dot{G},\ \partial_{A_7}G\}.
\]

Interpretation:
\begin{itemize}
    \item changes in group topology (coalitions, fragmentation),
    \item sensitivity to trust potentials and influence gradients,
    \item onset of coordination or polarisation,
    \item conditions for network bifurcations.
\end{itemize}


% ================================================================
\subsubsection{Jets of Roles \texorpdfstring{$R$}{R}}
% ================================================================

Roles formalise stable behavioural channels:
\[
\dot{R}=F_R(R,N,S,P).
\]

Jets:
\[
\partial_{A_7}R
=
\partial_{S}R\,\partial_{A_7}S
+\partial_{N}R\,\partial_{A_7}N
+\partial_{P}R\,\partial_{A_7}P.
\]

They encode:
\begin{itemize}
    \item role-switch dynamics,
    \item emergence of functional differentiation,
    \item instability when norms degrade,
    \item coupling between status and role coherence.
\end{itemize}


% ================================================================
\subsubsection{Jets of Status Distribution \texorpdfstring{$S$}{S}}
% ================================================================

Status field:
\[
\dot{S}=F_S(S,G,P,J_{\mathrm{influence}}).
\]

Jet:
\[
J^1(S)=\{\dot{S},\ \partial_{A_7}S\}.
\]

Interpretation:
\begin{itemize}
    \item status competition and stabilisation,
    \item influence propagation,
    \item transition from hierarchical to flat structures,
    \item bifurcations at institutional thresholds $\Theta_{\mathrm{inst}}$.
\end{itemize}


% ================================================================
\subsubsection{Jets of Norms and Institutional Constraints \texorpdfstring{$N$}{N}}
% ================================================================

Norm dynamics:
\[
\dot{N} = F_N(N,P,S,J_{\mathrm{sanction}}).
\]

Jets:
\[
\partial_{A_7}N 
= 
\partial_P N\, \partial_{A_7}P
+
\partial_S N\, \partial_{A_7}S.
\]

Norm jets capture:
\begin{itemize}
    \item stability of rule-sets,
    \item emergence of local conventions,
    \item institutional drift,
    \item conditions of breakdown when normative force falls below
          $\Theta_{\mathrm{norm}}$.
\end{itemize}


% ================================================================
\subsubsection{Jets of Social Potentials \texorpdfstring{$P_7$}{P_7}}
% ================================================================

Social potentials include trust, cohesion, sanction pressure, expectation
alignment, and cooperative incentive structure:
\[
P = (P_{\mathrm{trust}}, P_{\mathrm{cohesion}}, P_{\mathrm{sanction}},
P_{\mathrm{coord}}, P_{\mathrm{inst}}).
\]

Their jets:
\[
J^1(P)=\{\dot{P},\ \partial_{A_7}P\}.
\]

These derivatives quantify:
\begin{itemize}
    \item erosion or reinforcement of trust,
    \item formation of cooperative equilibria,
    \item phase transitions between coordination regimes,
    \item strengthening or weakening of institutions.
\end{itemize}


% ================================================================
\subsubsection{Jets of Social Flows \texorpdfstring{$J_7$}{J_7}}
% ================================================================

Flows in $K_7$ include:
\[
J_{\mathrm{comm}},\;
J_{\mathrm{resource}},\;
J_{\mathrm{influence}},\;
J_{\mathrm{sanction}},\;
J_{\mathrm{coord}}.
\]

Dynamics:
\[
\dot{J}_7 = F_J(J_7, P, G, S, N).
\]

Jet:
\[
J^1(J_7)=\{\dot{J}_7,\ \partial_{A_7}J_7\}.
\]

Interpretation:
\begin{itemize}
    \item stability of communication networks,
    \item propagation of influence and norms,
    \item coordination efficiency,
    \item onset of systemic conflict or alignment.
\end{itemize}


% ================================================================
\subsubsection{Jets of Social Thresholds \texorpdfstring{$\Theta_7$}{\Theta_7}}
% ================================================================

Key thresholds:
\[
\Theta_7 =
\{
\Theta_{\mathrm{trust}},\,
\Theta_{\mathrm{cohesion}},\,
\Theta_{\mathrm{norm}},\,
\Theta_{\mathrm{inst}},\,
\Theta_{\mathrm{coord}}
\}.
\]

Their jets:
\[
J^1(\Theta_7)=\{\dot{\Theta}_7,\ \partial_{A_7}\Theta_7\}.
\]

Threshold jets describe:
\begin{itemize}
    \item proximity to institutional collapse,
    \item formation of new stable norms,
    \item collective action failure,
    \item condensation of coordinated behaviour.
\end{itemize}


% ================================================================
\subsubsection{Jets of Social Cycles \texorpdfstring{$C_7$}{C_7}}
% ================================================================

The social continuum contains characteristic cycles:
\[
C_{\mathrm{coord}},\;
C_{\mathrm{norm}},\;
C_{\mathrm{status}},\;
C_{\mathrm{resource}},\;
C_{\mathrm{influence}},\;
C_{\mathrm{institution}}.
\]

A jet of a cycle $C_j$ is:
\[
J^1(C_j)
=
\{
\partial_t X_j,\;
\partial_{A_7}X_j
\},
\quad 
X_j\in\{G,R,S,N,P,J\}.
\]

Cycle jets diagnose:
\begin{itemize}
    \item coordination oscillations and coherence domains,
    \item status rivalry / stabilisation,
    \item normative reinforcement loops,
    \item resource redistribution dynamics,
    \item institutional drift and lock-in,
    \item conditions for emergence of $K_8$-level structures.
\end{itemize}


% ================================================================
\subsubsection{Jets and Boundary Stability \texorpdfstring{$\partial\Omega(K_7)$}{\partial\Omega(K_7)}}
% ================================================================

The social continuum becomes unstable when:
\[
\partial_{A_7}(P_{\mathrm{trust}} - \Theta_{\mathrm{trust}}) < 0,
\]
or
\[
\partial_{A_7}(P_{\mathrm{inst}} - \Theta_{\mathrm{inst}}) < 0,
\]
indicating breakdown of trust or institutional authority.

Similarly, coordination collapse occurs if
\[
\partial_{A_7}(P_{\mathrm{coord}} - \Theta_{\mathrm{coord}})<0.
\]

Jets quantify distances to these boundaries and predict which axes are
responsible for approaching $\partial\Omega$.


% ================================================================
\subsubsection{Jets and the Transition \texorpdfstring{$K_6 \to K_7$}{K_6 \to K_7}}
% ================================================================

The emergence of the social continuum arises when cognitive models
become aligned across agents:

\[
\partial_{A_7}G
\propto
\partial_{A_6}M_{\mathrm{shared}},
\qquad
\partial_{A_7}N
\propto
\partial_{A_6}\Pi_{\mathrm{collective}},
\]

\[
\partial_{A_7}P_{\mathrm{trust}}
\propto
\partial_{A_6}C_{\mathrm{model}}.
\]

Thus, jets of $K_6$ (model coherence, attention, prediction cycles)
stabilise and extend into the social domain, giving rise to coordinated
groups, norms, and institutions.


% ================================================================
\subsubsection{Summary}
% ================================================================

Jets on $K_7$ formalise the infinitesimal structure of social continua:

\begin{itemize}
    \item group topology and coordination,
    \item role and status differentiation,
    \item norm dynamics and institutional strength,
    \item social potentials (trust, cohesion, sanction),
    \item information and resource flows,
    \item social thresholds and boundary stability,
    \item characteristic coordination and norm cycles,
    \item emergence of collective structures from $K_6$ cognition.
\end{itemize}

They form the transition layer toward $K_8$, where symbolic and
civilisational structures emerge.
