% ================================================================
% ==== FILE: content/jets/jets_master.tex
% ================================================================

\section{Structural Jets}
\label{sec:jets-master}

The jet formalism introduced in this section provides a unified,
dimension-agnostic description of \emph{infinitesimal structural
propagation} across all continua in the Ontology of Continua (OC).
While the operators $E$, $\Psi$, $\Phi$ and $U$ describe macroscopic
evolution, embedding, deformation of admissible regions and global
continuumness, the jet hierarchy captures the \emph{local differential
structure} governing how continua deform along their axes, potentials,
flows, thresholds, and boundary gradients.

\medskip

In classical mathematics, jets represent truncated Taylor expansions of
smooth functions. In OC, jets generalise this idea: a jet is a local,
structurally admissible fragment of continuum evolution, encoding
\emph{all allowed derivatives of all structural components} up to a
specified order. This includes derivatives of:

\begin{itemize}
    \item axes $A(t)$ and their deformation modes,
    \item potentials $P(t)$ and their gradients $\partial P / \partial A$,
    \item thresholds $\Theta(t)$ and their rate of change,
    \item flows $J(t)$, including cross-level flows and boundary exchange,
    \item cycles $C(t)$ and their infinitesimal stability conditions,
    \item admissible regions $\Omega(K)$ and their boundary geometry $\partial\Omega(K)$,
    \item continuumness $k(t)$ and its differential response.
\end{itemize}

Thus, a jet represents the \emph{infinitesimal content} of the universal
evolution operator:
\[
E(K(t), M(t)) = K(t+\Delta t),
\]
providing a differential expansion that allows OC to connect macroscopic
evolution to local geometric constraints.



% ================================================================
\subsection{Definition of Structural Jets}
% ================================================================

For any continuum $K_x$, we define its $n$-th jet $J^n(K_x)$ as:

\begin{equation}
J^n(K_x)(t) = 
\left\{
\frac{d^m A}{dt^m},
\frac{d^m P}{dt^m},
\frac{d^m \Theta}{dt^m},
\frac{d^m J}{dt^m},
\frac{d^m C}{dt^m},
\frac{d^m \partial\Omega}{dt^m},
\frac{d^m k}{dt^m}
\ \middle|\ 
0 \le m \le n
\right\}.
\end{equation}

A jet exists if and only if:

\begin{enumerate}
    \item all derivatives up to order $n$ respect the admissibility
    conditions of the continuum;
    \item the deformation does not push $K_x$ outside $\Omega(M_x)$;
    \item thresholds at every order remain finite; and
    \item the structural tension at each order satisfies:
    \[
        T^{(m)}(t) < \Theta_{\mathrm{collapse}}^{(m)}.
    \]
\end{enumerate}

Thus, jets encode differential consistency conditions across all levels
of OC.



% ================================================================
\subsection{Role of Jets in Universal Dynamics}
% ================================================================

Jets serve four structural functions:

\paragraph{(1) Local generator of global evolution.}
While $E$ provides the global update, jets describe its infinitesimal
expansion:
\[
E(K(t)) = K(t) + dt \cdot J^1(K) + dt^2 \cdot J^2(K) + \cdots.
\]

\paragraph{(2) Compatibility conditions for \texorpdfstring{$\Psi$}{\Psi} and \texorpdfstring{$\Phi$}{\Phi}.}
The embedding operator $\Psi$ and the evolution of the ambient space
$\Phi$ must satisfy jet-consistency constraints, ensuring that:
\[
J^n(K_x) \subseteq J^n(M_x).
\]

\paragraph{(3) Local structure of phase transitions.}
Every dimensional transition (e.g., $K_{x} \to K_{x+1}$) begins as a
jet instability: an $n$th-order divergence along a new axis.

\paragraph{(4) Diagnosis of collapse modes.}
Death of a continuum corresponds to jet-level singularities:
\[
\left| \frac{d^m X}{dt^m} \right| \to \infty,
\]
for some structural component $X$.



% ================================================================
\subsection{Jets Across the Levels $K_0 \to K_{12}$}
% ================================================================

Every level $K_x$ has a characteristic jet structure:

\begin{itemize}
    \item $K_0$: trivial jet (no derivatives, no axes).
    \item $K_1$: geometric jets on $(X,\tau)$, first-order boundary motion.
    \item $K_2$: physical jets (fields, gradients, variational structure).
    \item $K_3$: chemical jets (activation energies, reaction profiles).
    \item $K_4$: biological jets (membrane curvature, gradient kinetics).
    \item $K_5$: electrical jets (ion flux derivatives, excitability).
    \item $K_6$: cognitive jets (prediction-error gradients).
    \item $K_7$: social jets (resource flows, institutional rigidity).
    \item $K_8$: civilizational jets (symbolic/technological propagation).
    \item $K_9$: theoretical jets (paradigm deformation, model flux).
    \item $K_{10}$: meta-theoretical jets (functorial consistency).
    \item $K_{11}$: hyper-theoretical jets (modal universe curvature).
    \item $K_{12}$: universal jets (derivatives of structural laws).
\end{itemize}

All specialised jet files extend this master definition.



% ================================================================
\subsection{Boundary and Threshold Constraints}
% ================================================================

For any jet to exist, the boundary and threshold conditions must remain
admissible:

\begin{align}
\partial\Omega(K_x)(t + dt) &= \partial\Omega(K_x)(t)
  + dt \cdot \frac{d}{dt}\partial\Omega(K_x)
  + O(dt^2), \\
\Theta_x(t + dt) &= \Theta_x(t)
  + dt \cdot \frac{d\Theta_x}{dt}
  + O(dt^2).
\end{align}

A jet exists if and only if:

\[
\partial\Omega(K_x)(t+dt) \subseteq \Omega(M_x)(t+dt),
\qquad
\Theta_x^{(m)} < \Theta_{\mathrm{max}}^{(m)}.
\]

Failures of these conditions correspond to structural collapse.



% ================================================================
\subsection{Summary}
% ================================================================

This master section establishes jets as the infinitesimal, level-agnostic
formalism linking all structural components of OC. It provides the
foundation for each specialised jets-file that follows, ensuring
consistency across $K_0 \to K_{12}$ and integration with the universal
operators $E$, $\Psi$, $\Phi$, and $U$.
