% ================================================================
% ==== FILE: content/jets/jets_k12.tex
% ================================================================

\section{Jets on $K_{12}$}
\label{sec:jets-k12}

Level $K_{12}$ represents the highest-order meta-ontological stratum:
a continuum whose states are admissible ontological universes,
together with their axes, potentials, thresholds, flows, cycles, and
evolution rules. Jets on $K_{12}$ capture infinitesimal changes in the
space of all possible continua, describing micro-perturbations in the
architecture of reality itself.


% ================================================================
\subsection{State Space of $K_{12}$}
% ================================================================

A state of $K_{12}$ is:
\[
\omega_{12} =
(\mathfrak{K},\ \mathfrak{A},\ \mathfrak{P},\ \mathfrak{\Theta},\ 
 \mathfrak{J},\ \mathfrak{C},\ \mathfrak{E}),
\]
where:
\begin{itemize}
    \item $\mathfrak{K}$ — admissible continua (the class of $K_0\!\to\!K_{11}$ structures),
    \item $\mathfrak{A}$ — allowed axes applicable to those continua,
    \item $\mathfrak{P}$ — meta-potentials governing entire ontological spaces,
    \item $\mathfrak{\Theta}$ — meta-thresholds determining possibility of continua,
    \item $\mathfrak{J}$ — admissible classes of flows,
    \item $\mathfrak{C}$ — allowed cycle types across continuum categories,
    \item $\mathfrak{E}$ — admissible evolution operators (e.g., $E$, $\Psi$, $\Phi$, $U$).
\end{itemize}

Jets on $K_{12}$ quantify infinitesimal variations of these meta-structures.


% ================================================================
\subsection{Jet Coordinates of $K_{12}$}
% ================================================================

\[
J^{m,n}(K_{12}) =
\left\{
\partial_t^k \partial_{\mathcal{X}}^\ell 
(\mathfrak{K},\mathfrak{A},\mathfrak{P},\mathfrak{\Theta},\mathfrak{J},\mathfrak{C},\mathfrak{E})
\right\},
\]
where:
\[
\mathcal{X} \in A_{12}
\]
is the **axis of meta-level variation**, controlling how the space of all admissible continua deforms.


% ================================================================
\subsection{Jets of Admissible Continua $\mathfrak{K}$}
% ================================================================

Dynamics:
\[
\dot{\mathfrak{K}}
=
F_{\mathfrak{K}}(\mathfrak{K},\mathfrak{A},\mathfrak{P},\mathfrak{\Theta}).
\]

Jet:
\[
J^1(\mathfrak{K})
= \{\dot{\mathfrak{K}},\ \partial_{A_{12}}\mathfrak{K}\}.
\]

Interpretation:
\begin{itemize}
    \item emergence of new admissible continuum types,
    \item extinction of unstable continuum families,
    \item deformation of allowable K-level hierarchies,
    \item restriction or expansion of existence conditions for $K_x$.
\end{itemize}


% ================================================================
\subsection{Jets of Meta-Axes $\mathfrak{A}$}
% ================================================================

Axes define which transformations are possible at any level.

Dynamics:
\[
\dot{\mathfrak{A}}
=
F_{\mathfrak{A}}(\mathfrak{K},\mathfrak{A},\mathfrak{\Theta}).
\]

Jet:
\[
J^1(\mathfrak{A})
= \{\dot{\mathfrak{A}},\ \partial_{A_{12}}\mathfrak{A}\}.
\]

Interpretation:
\begin{itemize}
    \item appearance of new ontological axes,
    \item removal of axes that violate meta-threshold conditions,
    \item shifts in the dimensional growth rules across continua.
\end{itemize}


% ================================================================
\subsection{Jets of Meta-Potentials $\mathfrak{P}$}
% ================================================================

Meta-potentials control the general "pressure landscape" shaping which
continua exist and how they evolve.

Dynamics:
\[
\dot{\mathfrak{P}}
=
F_{\mathfrak{P}}(\mathfrak{P},\mathfrak{\Theta},\mathfrak{K},\mathfrak{A}).
\]

Jet:
\[
J^1(\mathfrak{P})
= \{\dot{\mathfrak{P}},\ \partial_{A_{12}}\mathfrak{P}\}.
\]

Interpretation:
\begin{itemize}
    \item global shifts in meta-stability,
    \item modifications of inter-level compatibility,
    \item changes in the universal potential landscape for continua.
\end{itemize}


% ================================================================
\subsection{Jets of Meta-Thresholds $\mathfrak{\Theta}$}
% ================================================================

Meta-thresholds determine *which continua can exist at all*.

Dynamics:
\[
\dot{\mathfrak{\Theta}}
=
F_{\mathfrak{\Theta}}(\mathfrak{\Theta},\mathfrak{P},\mathfrak{K}).
\]

Jet:
\[
J^1(\mathfrak{\Theta})
=
\{\dot{\mathfrak{\Theta}},\ \partial_{A_{12}}\mathfrak{\Theta}\}.
\]

Interpretation:
\begin{itemize}
    \item tightening or loosening existence conditions for $K_x$,
    \item modification of the dimension-birth thresholds,
    \item global shifts in compatibility conditions between $K$-levels.
\end{itemize}


% ================================================================
\subsection{Jets of Meta-Flows $\mathfrak{J}$}
% ================================================================

Meta-flows represent permissible transformation patterns between
continua or between their structural components.

Dynamics:
\[
\dot{\mathfrak{J}}
=
F_{\mathfrak{J}}(\mathfrak{J},\mathfrak{K},\mathfrak{A},\mathfrak{P}).
\]

Jet:
\[
J^1(\mathfrak{J})
= \{\dot{\mathfrak{J}},\ \partial_{A_{12}}\mathfrak{J}\}.
\]

Interpretation:
\begin{itemize}
    \item birth of new universal transformation principles,
    \item suppression of flows causing cross-level inconsistency,
    \item emergence of novel meta-evolutionary patterns.
\end{itemize}


% ================================================================
\subsection{Jets of Meta-Cycles $\mathfrak{C}$}
% ================================================================

Cycles at $K_{12}$ encode stable patterns of universe-level structure.

Dynamics:
\[
\dot{\mathfrak{C}}
=
F_{\mathfrak{C}}(\mathfrak{C},\mathfrak{K},\mathfrak{A},\mathfrak{P},\mathfrak{\Theta}).
\]

Jet:
\[
J^1(\mathfrak{C})
= \{\dot{\mathfrak{C}},\ \partial_{A_{12}}\mathfrak{C}\}.
\]

Interpretation:
\begin{itemize}
    \item stability or fragility of meta-level recurrence,
    \item conditions for regeneration of continuum families,
    \item catastrophic breakage of universal structural cycles.
\end{itemize}


% ================================================================
\subsection{Jets of Meta-Evolution Operators $\mathfrak{E}$}
% ================================================================

Operators include $E$, $\Psi$, $\Phi$, $U$ and any future higher-order
evolution schemes.

Dynamics:
\[
\dot{\mathfrak{E}}
=
F_{\mathfrak{E}}(\mathfrak{E},\mathfrak{K},\mathfrak{A},\mathfrak{P},\mathfrak{\Theta}).
\]

Jet:
\[
J^1(\mathfrak{E})
= \{\dot{\mathfrak{E}},\ \partial_{A_{12}}\mathfrak{E}\}.
\]

Interpretation:
\begin{itemize}
    \item modifications of universe-level evolution laws,
    \item changes in operators that govern dimension birth or collapse,
    \item restructuring of compatibility conditions across ontological strata.
\end{itemize}


% ================================================================
\subsection{Jets and Boundary Geometry $\partial\Omega(K_{12})$}
% ================================================================

Approach to the boundary occurs when:

\[
\partial_{A_{12}}(\mathfrak{\Theta}_{\mathrm{exist}} - L_{\mathrm{min}}) > 0,
\]
\[
\partial_{A_{12}}(\mathfrak{P}_{\mathrm{global}} - P_{\mathrm{crit}}) > 0,
\]
\[
\partial_{A_{12}}(\mathrm{Compat}(K_x,K_y) - \Theta_{\mathrm{compat}}) > 0.
\]

Interpretation:
\begin{itemize}
    \item rising global instability of universe-level architecture,
    \item failure of cross-level compatibility conditions,
    \item collapse of entire families of continua.
\end{itemize}


% ================================================================
\subsection{Relation to $K_{11} \to K_{12}$ Transition}
% ================================================================

Transition occurs when:
\begin{itemize}
    \item formal ontologies ($K_{11}$) become variables of a higher
          meta-structure;
    \item thresholds and axes themselves become deformable entities;
    \item the evolution operator must operate on entire ontology classes,
          not singular structures;
    \item consistency becomes global rather than internal.
\end{itemize}

In jet form:
\[
\partial_{A_{12}}(\mathcal{O},\mathcal{R},\Delta,\Pi) 
\subset
\partial_{A_{12}}(\mathfrak{K},\mathfrak{A},\mathfrak{P},\mathfrak{\Theta}).
\]


% ================================================================
\subsection{Summary}
% ================================================================

Jets on $K_{12}$ describe:
\begin{itemize}
    \item infinitesimal variations in the space of possible universes;
    \item the birth or extinction of entire continuum families;
    \item global deformations of thresholds, potentials and axes;
    \item micro-dynamics of universe-level flows and cycles;
    \item evolution of evolution operators themselves;
    \item meta-stability, compatibility and catastrophic transitions.
\end{itemize}

They provide the differential structure required to understand how the
very possibility space of reality shifts, stabilises or collapses.
